### 2015-10-25 icfp-2015/paper.tex
Delete old XML samples and Erased type providers

\documentclass[preprint]{sigplanconf}

\usepackage{amssymb}
\usepackage{multicol}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage[fleqn]{amsmath}
\usepackage{graphics} 
\usepackage{stmaryrd}
\usepackage{amsthm}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage{ftnright}
\usepackage[T1]{fontenc}
\usepackage{semantic}
\usepackage{enumitem}
%\usepackage[hang,flushmargin]{footmisc} 
%\usepackage[notquote]{hanging}
%\usepackage{flushend}

% =================================================================================================

% TODO: Adapt XML example 
% TODO: Switch from sum types to variant types
% TODO: Discuss 'stability' of the inference (minor change in schema does not cause dramatic cahnges)
% TODO: Say that any variant is a subtype of any other variant 
% TODO: Say that anything is a subtype of variant
% TODO: Say that variant is a last-resort thing
% TODO: Double check that bottom/top are used correctly
% TODO: Explain (even more) that there is no least upper bound, which is why Theorem 3 is good enough
% TODO: Avoid "no other rule applies" in the formal system
% TODO: Practical evaluation (lots of downloads, lots of Github - how often people customize samples?)

\urlstyle{sf}

\makeatletter
\renewcommand{\@makefntext}[1]{%
  \parindent 1em%
  \raggedright
  \begin{hangparas}{0.8em}{1}
  \noindent {$^{\@thefnmark}$~#1}
  \end{hangparas}
}
\makeatother

\newcommand{\langl}{\begin{picture}(4.5,7)
\put(1.1,2.5){\rotatebox{60}{\line(1,0){5.5}}}
\put(1.1,2.5){\rotatebox{300}{\line(1,0){5.5}}}
\end{picture}}
\newcommand{\rangl}{\begin{picture}(4.5,7)
\put(.9,2.5){\rotatebox{120}{\line(1,0){5.5}}}
\put(.9,2.5){\rotatebox{240}{\line(1,0){5.5}}}
\end{picture}}

\newcommand{\lang}{\begin{picture}(5,7)
\put(1.1,2.5){\rotatebox{45}{\line(1,0){6.0}}}
\put(1.1,2.5){\rotatebox{315}{\line(1,0){6.0}}}
\end{picture}}
\newcommand{\rang}{\begin{picture}(5,7)
\put(.1,2.5){\rotatebox{135}{\line(1,0){6.0}}}
\put(.1,2.5){\rotatebox{225}{\line(1,0){6.0}}}
\end{picture}}

\newcommand{\llangl}{\langl\hspace{-0.35em}\langl}
\newcommand{\rrangl}{\rangl\hspace{-0.35em}\rangl}

\definecolor{cmtclr}{rgb}{0.0,0.6,0.0}
\definecolor{numclr}{rgb}{0.0,0.4,0.0}
\definecolor{kvdclr}{rgb}{0.0,0.0,0.6}
\definecolor{strclr}{rgb}{0.5,0.1,0.0}
\definecolor{prepclr}{rgb}{0.0,0.0,0.0}

\newcommand{\kvd}[1]{\textnormal{\textcolor{kvdclr}{\sffamily #1}}}
\newcommand{\num}[1]{\textnormal{\textcolor{numclr}{\sffamily #1}}}
\newcommand{\str}[1]{\textnormal{\textcolor{strclr}{\sffamily "#1"}}}
\newcommand{\strf}[1]{\textnormal{\textcolor{strclr}{\sffamily #1}}}
\newcommand{\ident}[1]{\textnormal{\sffamily #1}}
\newcommand{\lident}[1]{\textnormal{\sffamily\`{}\hspace{-0.25em}\`{}\hspace{-0.1em}#1\`{}\hspace{-0.25em}\`{}}}
\newcommand{\cmt}[1]{\textit{\sffamily\textcolor{cmtclr}{#1}}}

\newcommand{\lsep}[0]{\;\; | \;\;}
\newcommand{\narrow}[1]{\hspace{-0.85em} #1 \hspace{-0.85em}}

\newcommand{\tsep}[0]{\; \triangledown \;}
\newcommand{\tytag}{\ident{tag}}
\newcommand{\dropopt}[1]{\lfloor#1\rfloor}
\newcommand{\addopt}[1]{\lceil#1\rceil}
\newcommand{\tytagof}{\ident{tagof}}
\newcommand{\nameoftag}{\ident{nameof}}

\newcommand{\reduce}{\rightsquigarrow}

\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
\newcommand{\semalt}[1]{\llangl #1 \rrangl}

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

% =================================================================================================

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}
\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
\copyrightyear{20yy} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\doi{nnnnnnn.nnnnnnn}

\titlebanner{Unpublished draft, March 2015}        % These are ignored unless
%\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{F\# Data: \textnormal{Making structured data first-class citizens}}
%\subtitle{Subtitle Text, if any}

\authorinfo{Tomas Petricek}
           {University of Cambridge}
           {tomas@tomasp.net}
\authorinfo{Gustavo Guerra}
           {Microsoft Corporation, London}
           {gustavo@codebeside.org}
\authorinfo{Don Syme}
           {Microsoft Research, Cambridge}
           {dsyme@microsoft.com}
\maketitle

% =================================================================================================

\begin{abstract}
Most static type systems assume that a program runs in a closed world, but this is not the 
case. Modern applications interact with external services and often access data in structured 
formats such as XML, CSV and JSON. Static type systems do not understand such external data sources
and only make accessing data more cumbersome. Should we just give up and leave the messy world of 
external data to dynamic typing and runtime checks?

Of course, we should not give up! In this paper, we show how to integrate external data sources 
into the F\# type system. As most real-world data on the web do not come with an explicit schema, 
we develop a type inference algorithm that infers a type from representative samples. Next, we use 
the type provider mechanism for integrating the inferred structures into the F\# type
system.

The resulting library significantly reduces the amount of code that developers need to write when 
accessing data. It also provides additional safety guarantees. Arguably, as much safety as possible 
if we abandon the incorrect closed world assumption.
\end{abstract}

\category{D.3.3}{Programming Languages}{Language Constructs and Features }
\keywords F\#, Type Providers, JSON, XML, Data, Type Inference

% =================================================================================================
%
%   ###                                                                       
%    #  #    # ##### #####   ####  #####  #    #  ####  ##### #  ####  #    # 
%    #  ##   #   #   #    # #    # #    # #    # #    #   #   # #    # ##   # 
%    #  # #  #   #   #    # #    # #    # #    # #        #   # #    # # #  # 
%    #  #  # #   #   #####  #    # #    # #    # #        #   # #    # #  # # 
%    #  #   ##   #   #   #  #    # #    # #    # #    #   #   # #    # #   ## 
%   ### #    #   #   #    #  ####  #####   ####   ####    #   #  ####  #    # 
%
% =================================================================================================

\section{Introduction}
\label{sec:introduction}

The purpose of many modern applications is to connect multiple data sources or services and
present the aggregated data in a new form. Mobile applications for taking notes, searching train
schedules or finding tomorrow's weather all communicate with one or more services over the network.

Increasing number of such services provide REST-based end-points that return data as CSV, XML
or JSON. Despite numerous schematization efforts, the services typically do not come with an explicit 
schema. At best, the documentation provides a number of typical requests and sample responses.

For example, the OpenWeatherMap service provides an end-point for getting current weather for a 
given city\footnote{See ``Current weather data'': \url{http://openweathermap.org/current}}. The page documents the input 
URL parameters and shows one sample JSON response to illustrate the response structure.
Using a standard library for working with JSON and HTTP, we might call the service and 
read the temperature as follows\footnote{We abbreviate the full URL: 
\url{http://api.openweathermap.org/data/2.5/weather?q=Prague\&units=metric}}:

\noindent
\begin{equation*}
\begin{array}{l}
 \kvd{let}~\ident{data}=\ident{Http.Request}(\str{http://weather.org/?q=Prague}) \\[0.1em]
 \kvd{match}~\ident{JsonValue.Parse}(\ident{data})~\kvd{with} \\[0.1em]
 |~\ident{Record}(\ident{root})\rightarrow \\[0.1em]
 \quad \kvd{match}~\ident{Map.find}~\str{main}~\ident{root}~\kvd{with} \\[0.1em]
 \quad |~\ident{Record}(\ident{main})\rightarrow \\[0.1em]
 \quad \quad \kvd{match}~\ident{Map.find}~\str{temp}~\ident{main}~\kvd{with} \\[0.1em]
 \quad \quad |~\ident{Number}(\ident{num})\rightarrow \ident{printfn}~\str{Lovely \%f degrees!}~\ident{num} \\[0.1em]
 \quad \quad |~\_\rightarrow \ident{failwith}~\str{Incorrect format} \\[0.1em]
 \quad |~\_\rightarrow \ident{failwith}~\str{Incorrect format} \\[0.1em]
 |~\_\rightarrow \ident{failwith}~\str{Incorrect format} 
\end{array}
\end{equation*}
\vspace{0.1em}

\noindent
The code assumes that the response has a particular format described in the documentation. The
root node must be a record with a \str{main} field, which has to be another record containing
a \str{temp} field with a numerical value. When the format is incorrect, the data access simply fails
with an exception.

Using the JSON type provider from the F\# Data library, we can write code with exactly the 
same functionality in two lines:

\vspace{-0.6em}
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{W} = \ident{JsonProvider}\langl\str{http://weather.org/?q=Prague}\rangl \\[0.1em]
 \ident{printfn}~\str{Lovely \%f degrees!}~(\ident{W.GetSample().Main.Temp})
\end{array}
\end{equation*}
\vspace{0.0em}

\noindent
On the first line, $\ident{JsonProvider}\langl\str{...}\rangl$ invokes a type provider at 
compile-time with the URL as a sample. The type provider infers the structure of the response
and provides a type with a \ident{GetSample} method that returns a parsed JSON with nested
properties \ident{Main.Temp}, returning the temperature as a number.

In the rest of the paper, we give a detailed description of the mechanism and we also discuss
safety properties of the approach. The key novel contributions of this paper are:

\begin{itemize}
\item We present type providers for XML, CSV and JSON (Section~\ref{sec:providers}) 
  that are available in the F\# Data library and we also cover practical aspects of the 
  implementation that contributed to their industrial adoption (Section~\ref{sec:impl}).

\item We describe a predictable type inference algorithm for structured data formats that
  underlies the type providers (Section~\ref{sec:inference}). It is based on finding common 
  supertype of a set of examples.

\item We provide a minimal formal model (Section~\ref{sec:formal}) and use it to prove a
  relativized notion of type safety for our type providers (Section~\ref{sec:safety}). Although 
  we focus on our specific case, the relativized safety property demonstrates a new way of 
  thinking about ML-style type safety that is needed in the context of web.
\end{itemize}

\noindent
Although the F\# Data library \cite{fsharp-data} has been widely adopted and is one of the most downloaded F\# libraries 
it has not yet been presented in an academic form. Additional documentation and source code can 
be found at \url{http://fsharp.github.io/FSharp.Data}.

% =================================================================================================
%
%   #######                                                                                  
%      #    #   # #####  ######    #####  #####   ####  #    # # #####  ###### #####   ####  
%      #     # #  #    # #         #    # #    # #    # #    # # #    # #      #    # #      
%      #      #   #    # #####     #    # #    # #    # #    # # #    # #####  #    #  ####  
%      #      #   #####  #         #####  #####  #    # #    # # #    # #      #####       # 
%      #      #   #      #         #      #   #  #    #  #  #  # #    # #      #   #  #    # 
%      #      #   #      ######    #      #    #  ####    ##   # #####  ###### #    #  ####  
%
% =================================================================================================

\section{Structural type providers}
\label{sec:providers}

We start with an informal overview that shows how F\# Data type providers simplify working with 
JSON, XML and CSV. We also introduce the necessary aspects of F\# 3.0 type providers along the 
way. The examples in this section illustrate a number of key properties of our type inference 
algorithm:

\begin{itemize}
\item Our mechanism is robust and predictable. This is important as the user directly works with 
  the inferred types and should understand why a specific type was inferred from a given 
  sample\footnote{In particular, we do not use probabilistic methods where adding additional 
  sample could change the shape of the type.}.

\item Our inference mechanism prefers records over unions. This better supports developer tooling --
  most F\# editors provide code completion hints on ``.'' and so types with properties (records)
  are easier to use than types that require pattern matching.

\item Finally, we handle a number of practical concerns that may appear overly complicated, but
  are important in the real world. This includes support for different numerical types, \kvd{null}
  values and missing data and also different ways of representing Booleans in CSV files.
\end{itemize}

% --------------------------------------------------------------------------------------------------

\subsection{Working with JSON documents}
\label{sec:providers-json}

The JSON format used in the example in Section~\ref{sec:introduction} is a popular format for data
exchange on the web based on data structures used in JavaScript. The following is the definition 
of the \ident{JsonValue} type used earlier to represent parsed JSON:
%
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{JsonValue} = \\[0.1em]
 \quad|~\ident{Null} \\[0.1em]
 \quad|~\ident{Number}~\kvd{of}~\ident{decimal} \\[0.1em]
 \quad|~\ident{String}~\kvd{of}~\ident{string} \\[0.1em]
 \quad|~\ident{Boolean}~\kvd{of}~\ident{bool} \\[0.1em]
 \quad|~\ident{Record}~\kvd{of}~\ident{Map}\langl\ident{string}, \ident{JsonValue}\rangl \\[0.1em]
 \quad|~\ident{Array}~\kvd{of}~\ident{JsonValue}[] \\[0.1em]
\end{array}
\end{equation*}
%
The OpenWeatherMap example in the introduction used only a (nested) record containing a numerical
value. To demonstrate other aspects of the JSON type provider, we look at a more complex example
that also involves \kvd{null} value and an array:
%
{\small{
\begin{verbatim}
  [ { "name": "Jan", "age": 25 },
    { "name": "Alexander", "age": 3.5 },
    { "name": "Tomas" } ]
\end{verbatim}
}}
%
\noindent
Say we want to print the names of people in the list with an age if it is available. As before, 
the standard approach would be to pattern match on the parsed \ident{JsonValue}. The code would 
check that the top-level node is a \ident{Array}, iterate over the elements checking that each is
a \ident{Record} with certain properties and throw an exception or skip values in incorrect format.
The standard approach can be made nicer by defining helper functions. However, we still need to
specify names of fields as strings, which is error prone and can not be statically checked.

Assuming the \strf{people.json} file contains the above example and \ident{data} is a string value
that contains another data set in the same format, we can print names and ages as follows:
%
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{People}~=~\ident{JsonProvider}\langl\str{people.json}\rangl\hspace{1em} \\[0.6em]
 \kvd{let}~\ident{items}~=~\ident{People.Parse}(\ident{data})\\[0.1em]
 \kvd{for}~\ident{item}~\kvd{in}~\ident{items}~\kvd{do}\\[0.1em]
 \quad\ident{printf}~\str{\%s }~\ident{item.Name}\\[0.1em]
 \quad\ident{Option.iter}~(\ident{printf}~\str{(\%f)})~\ident{item.Age}
\end{array}
\end{equation*}
%
The code achieves the same simplicity as when using dynamically typed languages, but it is statically 
type-checked. In contrast to the earlier example, we now use a local file \strf{people.json} as a 
sample for the type inference, but then processes data from another source.

\paragraph{Type providers.}
The notation $\ident{JsonProvider}\langl\str{people.json}\rangl$ on the first line passes a 
\emph{static parameter} to the type provider. Static parameters are resolved at compile-time, 
so the file name has to be a constant. The provider analyzes the sample and generates
a type that we name \ident{People}. In editors that use the F\# Compiler Service, the 
type provider is also executed at development-time and so the same provided types are
used in code completion.

The \ident{JsonProvider} uses a type inference algorithm (Section~\ref{sec:inference})  and
infers the following types from the sample:
%
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{Entity}~=  \\[0.1em]
 \quad \kvd{member}~\ident{Name}~:~\ident{string} \\[0.1em]
 \quad \kvd{member}~\ident{Age}~:~\ident{option}\langl \ident{decimal}\rangl \\[0.6em]
 \kvd{type}~\ident{People}~=  \\[0.1em]
 \quad \kvd{member}~\ident{GetSample}~:~\ident{unit}~\rightarrow~\ident{Entity}[] \\[0.1em]
 \quad \kvd{member}~\ident{Parse}~:~\ident{string}~\rightarrow~\ident{Entity}[] \\[0.1em]
\end{array}
\end{equation*}
%
The type \ident{Entity} represents the person. The field \ident{Name} is available for all
sample values and is inferred as \ident{string}. The field \ident{Age} is marked as optional,
because the value is missing in one sample. The two sample ages are an integer $25$ and a 
decimal $3.5$ and so the common inferred type is \ident{decimal}.

The type \ident{People} provides two methods for reading data. \ident{GetSample} returns the
sample used for the inference and \ident{Parse} parses a JSON string containing data in the same 
format as the sample. Since the sample JSON is a collection of records, both methods return an 
array of \ident{Entity} values.

\paragraph{Error handling.}
In addition to the structure of the types, the type provider also specifies what code should be 
executed at run-time in place of \ident{item.Name} and other operations. This is done through a 
\emph{type erasure} mechanism demonstrated in Section~\ref{sec:providers-csv}. In this example,
the runtime behaviour is the same as in the hand-written sample in Section~\ref{sec:introduction} --
a member access throws an exception if the document does not have the expected format.

Informally, the safety property discussed in Section~\ref{sec:safety} states that if the inputs
are subtypes of some of the provided samples (\emph{i.e.}~the samples are representative), then
no exceptions will occur. In other words, we cannot avoid all failures, but we can prevent some.
If the OpenWeatherMap changes the response format, the sample in Section~\ref{sec:introduction}
will not re-compile and the user knows that the code needs to be changed. What this means for
the traditional ML type safety is discussed in Section~\ref{sec:safety-discuss}.

\paragraph{The role of records.}
The sample code is easy to write thanks to the fact that most F\# editors provide code completion
when ``.'' is typed. The developer does not need to look at the sample JSON file to see what fields
are available for a person. To support this scenario, our inference algorithm prefers records
(we also treat XML elements and CSV rows as records).

In the above example, this is demonstrated by the fact that \ident{Age} is marked as optional.
An alternative is to provide two different record types (one with \ident{Name} and other with 
\ident{Name} and \ident{Age}), but this would complicate the processing code.

% -------------------------------------------------------------------------------------------------

\subsection{Processing XML data, alternative}

In XML, documents are formed by (possibly) nested elements with attributes. We can view elements
as records (unlike in JSON, the records are named) with a field for each attribute and an additional
special field for the nested contents (which may be a collection of elements).

As an example, consider a simple extensible document format where a root element
{\ttfamily\small <doc>} can contain a number of document element, one of which is
{\ttfamily\small <heading>} representing headings:
%
{\small{
\begin{verbatim}
  <doc>
    <heading>Working with JSON</heading>
    <p>Type providers make this easy.</p>
    <heading>Working with XML</heading>
    <p>Processing XML is as easy as JSON.</p>
    <image source="xml.png" />
  </doc>
\end{verbatim}
}}
%
\noindent
The F\# Data library has been primarilly designed to allow easy reading of data. For example,
say we want to print all headings that appear in the document. The sample shows a part of the 
possible document structure (in particular the {\ttfamily\small <heading>} element we want to
read), but it does not show all possible elements (say, {\ttfamily\small <table>}). Assuming the
above document is \strf{sample.doc}, we can write:
%
\noindent
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{Document}~=~\ident{XmlProvider}\langl\str{sample.doc}\rangl\hspace{1em} \\[0.5em]
 \kvd{let}~\ident{root}~=~\ident{Document.Load}(\str{http://.../another.doc})\\
 \kvd{for}~\ident{elem}~\kvd{in}~\ident{root.Doc}~\kvd{do}\\
 \quad\kvd{match}~\ident{elem.Heading}~\kvd{with} \\
 \quad|~\ident{Some text} \rightarrow \ident{printf}~\str{ - \%s }~\ident{text}\\
 \quad|~\ident{None} \rightarrow ()\\
\end{array}
\end{equation*}
%
The example iterates over a collection of elements \ident{elem} returned by \ident{root.Doc}. 
Our system provides a typed access to elements known from the sample document and so we can
write \ident{elem.Heading}.

\paragraph{Open world and variant types.}
In the above example, The system does not infer a union type representing a choice between
heading, paragraph and image. By its nature, the XML format is extensible and the sample cannot 
include all possible examples\footnote{Even when the document structure is defined using XML 
Schema, documents may contain elements prefixed with other namespaces.}. The type of \ident{elem}
is thus a \emph{variant} type annotated with the three statically known possibilities (heading,
paragraph, image). The variant is mapped to the following F\# type:
%
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{Element}~=  \\[0.1em]
 \quad \kvd{member}~\ident{Heading}~:~\ident{option}\langl \ident{string} \rangl\\
 \quad \kvd{member}~\ident{Paragraph}~:~\ident{option}\langl \ident{string} \rangl\\
 \quad \kvd{member}~\ident{Image}~:~\ident{option}\langl \ident{Image} \rangl\\
\end{array}
\end{equation*}
%
The type provides a nice access to the elements known statically from the sample, but it encodes
the \emph{open-world assumption} -- the user also needs to handle the case when the value is none 
of the statically known elements.

One perhaps surprising consequence of the use of variant types is that the above code will work
correctly on document that contains elements other than those in the sample document. This would 
arguably not be desirable design for data types in a programming language, but it is extremely
useful when reading data from external data sources. It is also worth noting that our type 
inference attempts to minimze the number of variant types that appear in the resulting type --
a variant is not used if there is another option.

% --------------------------------------------------------------------------------------------------

\subsection{Reading CSV files}
\label{sec:providers-csv}

In the CSV file format, the structure is a collection of rows (records) consisting of fields 
(with names specified by the first row). The inference needs to infer the types of fields.
For example:
%
{\small{
\begin{verbatim}
  Ozone, Temp, Date,       Autofilled
  41,    67,   2012-05-01, 0
  36.3,  72,   2012-05-02, 1
  12.1,  74,   3 kveten,   0
  17.5,  #N/A, 2012-05-04, 0
\end{verbatim}
}}
%
\noindent
One difference between JSON and CSV formats is that in CSV, the literals have no data types.
In JSON, strings are \str{quoted} and Booleans are \kvd{true} and \kvd{false}. This is not the
case for CSV and so we need to infer not just the structure, but also the primitive types.

Assuming the sample is saved as \strf{airdata.csv}, the following snippet prints all
rows from another file that were not autofilled:
%
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{AirCsv}~=~\ident{CsvProvider}\langl\str{airdata.csv}\rangl\hspace{1em} \\[0.6em]
 \kvd{let}~\ident{air}~=~\ident{AirCsv.Parse}(\ident{data})\\[0.1em]
 \kvd{for}~\ident{row}~\kvd{in}~\ident{air.Rows}~\kvd{do}\\[0.1em]
 \quad\kvd{if}~\ident{not}~\ident{row.Autofilled}~\kvd{then}\\[0.1em]
 \quad\quad \ident{printf}~\str{\%s: \%d}~\ident{row.Date row.Ozone}\\
\end{array}
\end{equation*}
%
The type of the record (\ident{row}) is, again, inferred from the sample. The \ident{Date} column
uses mixed formats and is inferred as \ident{string} (although we support many date formats and 
``May 3'' would work). More interestingly, we also infer \ident{Autofilled} as Boolean, because 
the sample contains only $0$ and $1$ values and using those for Booleans is a common CSV convention.
Finally, the fields \ident{Ozone} and \ident{Temp} have types \ident{decimal} and
$\ident{option}\langl\ident{int}\rangl$.

% -------------------------------------------------------------------------------------------------

\subsection{Summary}
The previous three sections demonstrated the F\# Data type providers for JSON, CSV and XML.
A reader interested in more examples is invited to look at documentation on the F\# Data web 
site\footnote{Available at \url{http://fsharp.github.io/FSharp.Data/}}.

It should be noted that we did not attempt to present the library using well-structured ideal 
samples, but instead used data sets that demonstrate typical issues that are frequent in real 
world inputs (missing data, inconsistent encoding of Booleans and heterogeneous structures).

From looking at just the previous three examples, these may appear arbitrary, but our experience
suggests that they are the most common issues. The following JSON response with government debt
information returned by the World Bank API demonstrates all three problems:
%
{\small{
\begin{verbatim}
  [ { "page": 1, "pages": 5 },
    [ { "indicator": "GC.DOD.TOTL.GD.ZS",
        "date": "2012",
        "value": null },
      { "indicator": "GC.DOD.TOTL.GD.ZS",
        "date": "2010",
        "value": "35.1422970266502" } ] ]
\end{verbatim}
}}
%
\noindent
First of all, the top-level element is a collection containing two values of different kind.
The first is a record with meta-data about the current page and the second is an array with data. 
The JSON type provider infers this as heterogeneous collection with properties \ident{Record}
for the former and \ident{Array} for the latter. Second, the \ident{value} is \kvd{null} for some
records. Third, numbers can be represented in JSON as numeric literals (without quotes), but
here, they are returned as string literals instead\footnote{This is often used to avoid non-standard
numerical types in JavaScript.}.

% =================================================================================================
%
%   #######                                                                                  
%      #    #   # #####  ######    # #    # ###### ###### #####  ###### #    #  ####  ###### 
%      #     # #  #    # #         # ##   # #      #      #    # #      ##   # #    # #      
%      #      #   #    # #####     # # #  # #####  #####  #    # #####  # #  # #      #####  
%      #      #   #####  #         # #  # # #      #      #####  #      #  # # #      #      
%      #      #   #      #         # #   ## #      #      #   #  #      #   ## #    # #      
%      #      #   #      ######    # #    # #      ###### #    # ###### #    #  ####  ###### 
%
% =================================================================================================

\section{Structural type inference}
\label{sec:inference}

Our type inference algorithm for structured formats is based on a subtyping relation. When 
inferring the type of a document, we infer the most specific types of individual values (CSV rows,
JSON or XML nodes) and find a common supertype of values in the sample.

In this section, we define the \emph{structural type} $\sigma$, which is a type of structured data.
Note that this type is distinct from the programming language types $\tau$ (the type generates the
latter from the former). Next, we define the subtyping relation on structural types $\sigma$ and 
describe the algorithm for finding a common supertype.

Note that the subtyping relation between structural types $\sigma$ does not map to a subtyping 
relation between F\#  types $\tau$. However, it does hold at runtime, meaning that an operation on 
a supertype can be performed on its subtypes. We make the type provider mechanism and runtime aspects 
precise in Section~\ref{sec:formal} before discussing safety in Section~\ref{sec:safety}.

\subsection{Structural types}
\label{sec:inference-types}

The grammar below defines \emph{structural type} $\sigma$. We distinguish between \emph{non-nullable types}
that always have a valid value (written as $\hat{\sigma}$) and \emph{nullable types} that encompass missing 
and \kvd{null} values (written as $\sigma$). We write $\nu$ for record field names and $\nu?$
for record names, which can be empty:
%
\begin{equation*}
\begin{array}{rcl}
 \hat{\sigma} &\narrow{=}& \nu?\; \{ \nu_1 : \sigma_1, \ldots, \nu_n : \sigma_n \} \\[0.1em]
                &\narrow{|}& \ident{float} \lsep \ident{decimal} \lsep \ident{int} \lsep \ident{bit} \lsep \ident{bool} \lsep \ident{string} 
 \\[0.6em] 
       \sigma &\narrow{=}& ~\hat{\sigma}~ \lsep \hat{\sigma}\;\,\kvd{option} \lsep [\sigma] \\[0.1em]
              &\narrow{|}& \sigma_1 + \ldots + \sigma_n \lsep ~\top~ \lsep \kvd{null}
\end{array}
\end{equation*}
%
The non-nullable types include records (consisting of an optional name and zero or more fields with
their types) and primitive types. Records arising from XML documents are always named, while records
used by JSON and CSV providers are always unnamed.

Next, we have three numerical types -- \ident{int} for integers, \ident{decimal} for small high-precision 
decimal numbers and \ident{float} for floating-point numbers (these are related by sub-typing relation as
discussed in the next section). Furthermore, the \ident{bit} type is a type of values $0$ and $1$ and 
makes it possible to infer Boolean in the CSV processing example discussed in Section~\ref{sec:providers-csv}.

Any non-nullable type is also a nullable type, but it can be wrapped in the \kvd{option}
constructor to explicitly permit the \kvd{null} value. These are typically mapped to the standard F\# option 
type. A simple collection type $[\sigma]$ is also nullable and missing values or \kvd{null} are treated as 
empty collection. The type $\kvd{null}$ is inhabited by the $\kvd{null}$ value (using an overloaded but not
ambiguous notation) and $\top$ represents the top type.

Finally, a union type in our model implicitly permits the \kvd{null} value. This is because structural union 
types are \emph{not} mapped to F\# union types. Instead our encoding requires the user to handle the 
cases one-by-one and so they also need to handle the situation when none of the cases matches (as discussed
in Section~\ref{sec:impl-unions}, this is the right choice for an F\# type provider, but it would not 
necessarily be a good fit for other languages).

% TODO: Also include reference to section with more on heterogeneous collections.

% --------------------------------------------------------------------------------------------------

\subsection{Subtyping relation}
\label{sec:inference-subtyping}

To provide a basic intuition, the subtyping relation between structural types is illustrated in
Figure~\ref{fig:subtyping-diagram}. We split the diagram in two parts. The upper part shows 
non-nullable types (with records and primitive types). The lower part shows nullable types with 
\kvd{null}, collections and optional values. We omit links between the two part, but any type
$\hat{\sigma}$ is a subtype of $\hat{\sigma}~\kvd{option}$ (in the diagram, we abbreviate 
$\sigma~\kvd{option}$ as $\sigma?$).

The diagram shows only the basic structure and it does not explain relationships between records
with the same name and between union types. This following definition makes this precise.

\begin{definition}
We write $\sigma_1 :> \sigma_2$ to denote that $\sigma_2$ is a subtype of $\sigma_1$. The 
subtyping relation is defined as a transitive reflexive closure of these rules:

\vspace{0.5em}
\noindent\boxed{\emph{Primitives, options, collections}}
\begin{align}
\label{eq:sub-prim}\tag{B1}
\ident{float}\,:>\,\ident{decimal}\,:>\,\ident{int}\,:>\,\ident{bit}\quad&\textnormal{and}\quad\ident{bool}\,:>\,\ident{bit}\\[-0.2em]
\label{eq:sub-null}\tag{B2}
\sigma :> \kvd{null} \quad &(\textnormal{iff}~\sigma \neq \hat{\sigma}) \\[-0.2em]
\label{eq:sub-opt}\tag{B3}
\hat{\sigma}~\kvd{option} :> \hat{\sigma} \quad &(\textnormal{for all}~\hat{\sigma})\\[-0.2em]
\label{eq:sub-top}\tag{B4}
\sigma :> \top \quad &(\textnormal{for all}~\sigma)\\[-0.2em]
\label{eq:sub-col}\tag{B5}
[\sigma_1] :> [\sigma_2] \quad &(\textnormal{if}~\sigma_1 :> \sigma_2)
\end{align}

\noindent\boxed{\emph{Union types}}
\begin{align}
\label{eq:sub-union1}\tag{U1}
\sigma_1 + \ldots + \sigma_n :> \sigma_1' + \ldots + \sigma_n' \quad &(\textnormal{iff}~\forall i.\sigma_i :> \sigma_i') \\
\label{eq:sub-union2}\tag{U2}
\sigma_1 + \ldots + \sigma_n :> \sigma_1 + \ldots + \sigma_m \quad &(\textnormal{where}~m \leq n) \\
\label{eq:sub-union3}\tag{U3}
\sigma_1 + \ldots + \sigma_n :> \sigma_1 + \ldots + \sigma_m \quad &(\textnormal{where}~m \geq n) \\
\label{eq:sub-union4}\tag{U4}
\hspace{-1em}\sigma_1 + \ldots + \sigma_n :> \sigma_{\pi(1)} + \ldots + \sigma_{\pi(n)} \quad &(\textnormal{permutation}~\pi) \\
\label{eq:sub-union-el}\tag{U5}
\sigma_1 + \ldots + \sigma_n :> \sigma_i \quad &(\textnormal{for}~i \in 1 \ldots n)
\end{align}

\noindent\boxed{\emph{Record types}}
\begin{align}
\hspace{-1em}
\label{eq:sub-record1}\tag{R1}
 \nu?~\{ \nu_1\!:\!\sigma_1, .., \nu_n\!:\!\sigma_n \} \hspace{-0.1em}:> \hspace{-0.1em}
 \nu?~\{ \nu_1\!:\!\sigma_1', .., \nu_n\!:\!\sigma_n' \}
 \; (\sigma_i \hspace{-0.25em}:>\hspace{-0.25em} \sigma_i')
\end{align}
\vspace{-1.4em}
\begin{align}
\hspace{-1em}
\label{eq:sub-record2}\tag{R2}
 \nu?~\{ \nu_1\!:\!\sigma_1, .., \nu_n\!:\!\sigma_n \} \hspace{-0.1em} :> \hspace{-0.1em}
 \nu?~\{ \nu_1\!:\!\sigma_1, .., \nu_m\!:\!\sigma_m \}
 \; (m \hspace{-0.1em}\geq\hspace{-0.1em} n)
\end{align}
\vspace{-1.0em}
\begin{align}
\label{eq:sub-record3}\tag{R3}
\hspace{-1.75em}
\begin{array}{l}
 \nu?~\{ \nu_1\!:\!\sigma_1, .., \nu_n\!:\!\sigma_n \} :> \\
 \nu?~\{ \nu_{\pi(1)}\!:\!\sigma_{\pi(1)}, .., \nu_{\pi(m)}\!:\!\sigma_{\pi(m)} \}
\end{array} \qquad &(\textnormal{permutation}~\pi)
\end{align}
\vspace{-0.6em}
\begin{align}
\label{eq:sub-record4}\tag{R4}
\hspace{-1.75em}
\begin{array}{l}
 \nu?~\{ \nu_1\!:\!\sigma_1, .., \nu_n\!:\!\sigma_n, \nu_{n+1}\!:\!\kvd{null}, .., \nu_{n+m}\!:\!\kvd{null} \} :> \\
 \nu?~\{ \nu_1\!:\!\sigma_1, .., \nu_n\!:\!\sigma_n \}
\end{array}
\end{align}
\end{definition}

% --------------------------------------------------------------------------------------------------

\begin{figure}
\begin{center}
\includegraphics[scale=0.75,trim=5mm 5mm 5mm 5mm,clip]{images/hierarchy.pdf} % left bottom right top
\end{center}
\vspace{-0.5em}
\caption{Subtype relation between structural types}
\label{fig:subtyping-diagram}
\vspace{-0.5em}
\end{figure}

% -------------------------------------------------------------------------------------------------

\begin{figure*}[t]
\noindent
\begin{equation*}
\hspace{3em}
\inference[(record-1)\;]
  { (\nu_i = \nu'_j \Leftrightarrow (i = j) \wedge (i \leq k))
      \qquad \forall i\in\{ 1 .. k \}.(\sigma_i \tsep \sigma'_i \vdash \sigma''_i) }
  { \begin{array}{l}
    \nu? \; \{ \nu_1 : \sigma_1,  \; \ldots \;, \nu_k : \sigma_k, \; \ldots \;, \nu_n : \tau_n \} \tsep
    \nu? \; \{ \nu'_1 : \sigma'_1, \; \ldots \;, \nu'_k : \sigma'_k, \; \ldots \;, \nu'_m : \tau'_m \} \vdash\\
    \nu? \; \{ \nu_1 : \sigma''_1, \; \ldots \; , \nu_k : \sigma''_k, 
                            \nu_{k+1} : \addopt{\sigma_{k+1}}, \ldots, \nu_n : \addopt{\sigma_n},
                            \nu'_{k+1} : \addopt{\sigma'_{k+1}}, \ldots, \nu'_m : \addopt{\sigma'_m} \}
    \end{array} }
\end{equation*}

% union with something else
\begin{equation*}
\textnormal{\footnotesize{(union-1)}}\;\;
\inference
  {\exists i . \tytagof(\sigma_i) = \tytagof(\sigma) & \sigma \tsep \sigma_i \vdash \sigma_i' & \tytagof(\sigma)\neq\ident{union}}
  {\sigma \tsep (\sigma_1 + \ldots + \sigma_n) \vdash (\sigma_1 + \ldots + \dropopt{\sigma_i'} + \ldots + \sigma_n)}
\quad\quad
\inference
  {\nexists i . \tytagof(\sigma_i) = \tytagof(\sigma) & \tytagof(\sigma)\neq\ident{union}}
  {\sigma \tsep (\sigma_1 + \ldots + \sigma_n) \vdash (\sigma_1 + \ldots + \sigma_n + \dropopt{\sigma})}
\end{equation*}

% two union types
\begin{equation*}
\inference[(union-2)\;]
  { \tytagof(\sigma_i) = \tytagof(\sigma'_i) \quad 
    \sigma_i \tsep \sigma_i' \vdash \sigma''_i \qquad (\forall i\in \{ 1 .. k \}) }
  { \begin{array}{l}
    (\sigma_1  + \ldots + \sigma_k +  \ldots + \sigma_n) \tsep 
    (\sigma'_1 + \ldots + \sigma'_k + \ldots + \sigma'_m) \vdash\\
    (\sigma''_1 + \ldots + \sigma''_k + \sigma_{k+1} + \ldots + \sigma_{n} + \sigma'_{k+1} + \ldots + \sigma'_{m}) 
    \end{array} }
%
\qquad\quad
\inference[(union-3)\;]
  {(\textnormal{no other rule applies})}
  {\sigma_1 \tsep \sigma_2 \vdash \dropopt{\sigma_1} + \dropopt{\sigma_2}}
\end{equation*}

\begin{equation*}
\hspace{1em}
\inference[(order-1)\;]
  { \nu? \; \{ \nu_1 : \sigma_1, \ldots, \nu_n : \sigma_n \} \tsep \sigma \vdash \sigma' }
  { \nu? \; \{ \nu_{\pi(1)} : \sigma_{\pi(1)}, \ldots, \nu_{\pi(n)} : \sigma_{\pi(n)} \} \tsep \sigma \vdash \sigma' }
\quad
\inference[(order-2)\;]
  { \sigma_1 + \ldots + \sigma_n \tsep \sigma \vdash \sigma' }
  { \sigma_{\pi(1)} + \ldots + \sigma_{\pi(n)} \tsep \sigma \vdash \sigma' }
\qquad (\pi~\textnormal{permutation})  
\end{equation*}

\begin{equation*}
\hspace{3em}
% collection types
\inference[(list)\;]
  {\sigma_1 \tsep \sigma_2 \vdash \sigma}
  {[\sigma_1] \tsep [\sigma_2] \vdash [\sigma]}
\qquad\qquad
% primitive types
\inference[(prim)\;]
  {\sigma_1 :> \sigma_2}
  {\sigma_1 \tsep \sigma_2 \vdash \sigma_1}\quad
\begin{array}{l}
  (\textnormal{when}\,~\sigma_1, \sigma_2 \in \{\ident{bit}, \ident{int}, \ident{decimal}, \ident{float} \} \\
   \quad~~\;\textnormal{or}\,~\sigma_1, \sigma_2 \in \{\ident{bit}, \ident{bool} \})
\end{array}   
\end{equation*}

% all rules are symmetric and reflexive
\noindent
\begin{equation*}
\hspace{4.75em}
\inference[(sym)\;]
  {\sigma_1 \tsep \sigma_2 \vdash \sigma}
  {\sigma_2 \tsep \sigma_1 \vdash \sigma}
%
\qquad\quad
\begin{array}{l}
 \textnormal{\footnotesize{(refl)}}\;\; \sigma \tsep \sigma \vdash \sigma\\[0.6em]
 % top type
 \textnormal{\footnotesize{(top)}}\;\; \top \tsep \sigma \vdash \sigma
\end{array}
%
\qquad\quad 
% null and nullable
 \begin{array}{l}
 \textnormal{\footnotesize{(null-1)}}\;\; \sigma \tsep \kvd{null} \vdash \sigma \quad(\sigma :> \kvd{null}) \\[0.6em]
 \textnormal{\footnotesize{(null-2)}}\;\; \sigma \tsep \kvd{null} \vdash \sigma~\kvd{option} \quad(\sigma :\ngtr \kvd{null})
 \end{array}
\end{equation*}

\caption{Inference judgements that define the common supertype relation}
\label{fig:subtyping-cst}
\end{figure*}

% --------------------------------------------------------------------------------------------------

\noindent
Here is a summary of the key aspects of the definition:
\begin{itemize}
\item For numeric types (\ref{eq:sub-prim}), we infer a single most precise numeric type that 
  can represent all values from a sample dataset, so we order types by the ranges they represent;
  \ident{int} is a 32-bit integer, \ident{decimal} has a range of about $-1^{29}$ to $1^{29}$ and 
  \ident{float} is a double-precision floating-point.

\item The \ident{bit} type is a subtype of both \ident{int} and \ident{bool} (\ref{eq:sub-prim}). 
  Thus, a sample $0,1$ has a type \ident{bit}; $0,1,\kvd{true}$ is \ident{bool} and
  $0,1,2$ becomes \ident{int} (for a sample $0,1,2, \kvd{true}$ we would need a union type).

\item The \kvd{null} type is a subtype of all nullable types (\ref{eq:sub-null}), that is all 
  $\sigma$ types excluding non-nullable types $\hat{\sigma}$). Any non-nullable type is also a 
  subtype of its optional version (\ref{eq:sub-opt}).

\item There is a top type (\ref{eq:sub-top}), but no bottom type. It is possible to find common 
  supertype of any two types, because a union type $\tau_1 + \ldots + \tau_n$ is a supertype of all its 
  components (\ref{eq:sub-union-el}).

\item For unions, we include the standard rules. Subtype can have fewer alternatives 
  (\ref{eq:sub-union1}), elements can be reordered (\ref{eq:sub-union2}) and subtyping is covariant 
  (\ref{eq:sub-union4}). An unusual rule (\ref{eq:sub-union3}) says that subtype can also have 
  \emph{more} cases. This is valid because the type provider exposes unions as objects that do not
  allow complete pattern matching.

\item As usual, the subtyping on records is covariant (\ref{eq:sub-record1}), subtype can have additional 
  fields (\ref{eq:sub-record2}) and fields can  be reordered (\ref{eq:sub-record3}). The interesting rule 
  is the last one (\ref{eq:sub-record4}).  Together with covariance, it states that a subtype can omit some 
  fields, provided that their types are nullable.
\end{itemize}

\noindent
The rule that allows subtype to have fewer record elements (\ref{eq:sub-record4}) is particularly
important. It allows us to prefer records in some cases. For example, given two samples 
$\{ \ident{name}:\ident{string} \}$ and $\{ \ident{name}:\ident{string}, \ident{age}:\ident{int} \}$,
we can find a common supertype $\{ \ident{name}:\ident{string}, \ident{age}:\ident{int}~\kvd{option} \}$
which is also a record. For usability reasons, we prefer this to another common supertype
$\{ \ident{name}:\ident{string} \} + \{ \ident{name}:\ident{string}, \ident{age}:\ident{int} \}$.
The next section describes precisely how our inference algorithm works.

It is worth noting that some of the aspects (such as 3 different numeric types and handling of 
the \kvd{null} type) are specific to F\#. Similar problems will appear with most real-world
languages and systems. JVM and OCaml all haves multiple numeric types, but they would handle \kvd{null}
differently.

We could also be more or less strict about missing data -- we choose to handle missing values 
silently when we can (\kvd{null} becomes an empty collection), but we are explicit in other cases 
(we infer \ident{string}~\kvd{option} as a hint to the user rather than treating \kvd{null} as an 
empty string). These choices are based on experience with using F\# Data in practice.

% -------------------------------------------------------------------------------------------------

\begin{figure*}
\begin{equation*}
\begin{array}{l}
 \ident{asInt}(i) \,\reduce i\\
 \ident{asInt}(d) \reduce i\quad(i = \lfloor d \rfloor)\\
 \ident{asInt}(f) \reduce i\quad(i = \lfloor f \rfloor) \\
 \ident{asInt}(t) \,\reduce i\quad(t~\text{\small represents}~i) \\
 \ident{asInt}(\kvd{true}) \reduce 1\\
 \ident{asInt}(\kvd{false}) \reduce 0\\
\end{array}
\;
\begin{array}{l}
 \ident{asDec}(i) \,\reduce d\quad(d = i)\\
 \ident{asDec}(d) \reduce d\\
 \ident{asDec}(f) \reduce d\quad(d = \lfloor f \rfloor) \\
 \ident{asDec}(t) \,\reduce d\quad(t~\text{\small represents}~d) \\
 \ident{asDec}(\kvd{true}) \reduce 1.0\\
 \ident{asDec}(\kvd{false}) \reduce 0.0\\
\end{array}
\;
\begin{array}{l}
 \ident{asFloat}(i) \,\reduce f\quad(f = i)\\
 \ident{asFloat}(d) \reduce f\quad(f = d)\\
 \ident{asFloat}(f) \reduce f\\
 \ident{asFloat}(t) \,\reduce f\quad(t~\text{\small represents}~f) \\
 \ident{asFloat}(\kvd{true}) \reduce 1.0\\
 \ident{asFloat}(\kvd{false}) \reduce 0.0\\
\end{array}
\;
\begin{array}{l}
 \ident{asStr}(i) \,\reduce t\quad(t~\text{\small represents}~i)\\
 \ident{asStr}(d) \reduce t\quad(t~\text{\small represents}~d)\\
 \ident{asStr}(f) \reduce t\quad(t~\text{\small represents}~f)\\
 \ident{asStr}(t) \,\reduce t\\
 \ident{asStr}(\kvd{true}) \reduce \str{true}\\
 \ident{asStr}(\kvd{false}) \reduce \str{false}\\
\end{array}
\end{equation*}
%
\begin{equation*}
\hspace{-1.5em}
\begin{array}{l}
 \ident{asBool}(1) \reduce \kvd{true}\\
 \ident{asBool}(1.0) \reduce \kvd{true}\\
 \ident{asBool}(\str{true}) \reduce \kvd{true}\\
 \ident{asBool}(\str{1}) \reduce \kvd{true}
 \\[0.6em]
 \ident{isNull}(\kvd{null}) \reduce \kvd{true} \\
 \ident{isNull}(\_) \reduce \kvd{false} 
\end{array}
\hspace{2.6em}
\begin{array}{l}
 \ident{asBool}(0) \reduce \kvd{false}\\
 \ident{asBool}(0.0) \reduce \kvd{false}\\
 \ident{asBool}(\str{false}) \reduce \kvd{false}\\
 \ident{asBool}(\str{0}) \reduce \kvd{false}\\
 \ident{asBool}(b) \reduce b
 \\[0.6em]
 \ident{getChildren}(\kvd{null}) \reduce [~]
\end{array}
\hspace{3.0em}
\begin{array}{l}
 \ident{getField}(\nu,\nu_i, \nu~\{\ldots, \nu_i=s_i, \ldots\}) \reduce s_i\\
 \ident{getField}(\bot, \nu_i, \{\ldots, \nu_i=s_i, \ldots\}) \reduce s_i\\
 \ident{getField}(\nu,\nu', \nu~\{\ldots, \nu_i=s_i, \ldots\}) \reduce \kvd{null}\quad(\nexists i.\nu_i=\nu' )\\
 \ident{getField}(\bot, \nu', \{\ldots, \nu_i=s_i, \ldots\}) \reduce \kvd{null}\quad(\nexists i.\nu_i=\nu' )
 \\[0.6em]
 \ident{getChildren}([s_1; \ldots; s_n]) \reduce [s_1; \ldots; s_n]
\end{array}
\end{equation*}
%
\begin{equation*}
\hspace{-1.5em}
\begin{array}{l}
 \ident{isTag}(\ident{string}, t) \reduce \kvd{true} \\
 \ident{isTag}(\ident{bool}, v) \reduce \kvd{true} \quad(\textnormal{when}~v\in{0,1,0.0, 1.0, \kvd{true},\kvd{false}} )\\
 \ident{isTag}(\ident{number}, v) \reduce \kvd{true} \quad(\textnormal{when}~v=i, v=d, v=f) \\
 \ident{isTag}(\ident{collection}, [s_1; \ldots; s_n]) \reduce \kvd{true} \\
\end{array}
\hspace{1.8em}
\begin{array}{l}
 \ident{isTag}(\ident{collection}, \kvd{null}) \reduce \kvd{true} \\
 \ident{isTag}(\ident{rec-anon}, \{ \nu_1\mapsto s_1, \ldots, \nu_n\mapsto s_n \}) \reduce \kvd{true} \\
 \ident{isTag}(\ident{rec-named}~\nu, \nu\;\{ \nu_1\mapsto s_1, \ldots, \nu_n\mapsto s_n \}) \reduce \kvd{true} \\
 \ident{isTag}(\_, \_) \reduce \kvd{false} \\
\end{array}
\end{equation*}

\caption{Reduction rules for conversion functions}
\label{fig:op-conversions}
\end{figure*}

% -------------------------------------------------------------------------------------------------

\subsection{Common supertype relation}
\label{sec:inference-commonsuper}

As mentioned earlier, the structural type inference relies on a finding common supertype. However,
the partially ordered set of types does not have a unique greatest lower bound. For example, record types
$\{\ident{a}:\ident{int}\}$ and $\{\ident{b}:\ident{bool}\}$ have common supertypes
$\{\ident{?a}:\ident{int}, \ident{?b}:\ident{bool}\}$ and $\{\ident{a}:\ident{int}\} + \{\ident{b}:\ident{bool}\}$
which are unrelated. Our inference algorithm always prefers records over unions (Theorem~\ref{thm:no-unions}). 
The definition of the \emph{common supertype} relation uses a number of auxiliary definitions 
that are explained in the discussion that follows.

\begin{definition}
A \emph{common supertype} of types $\sigma_1$ and $\sigma_2$ is a type $\sigma$, written 
$\sigma_1 \triangledown \sigma_2 \vdash \sigma$, obtained according to the inference rules in 
Figure~\ref{fig:subtyping-cst}.
\end{definition}

\noindent
When finding a common supertype of two records (\emph{record-1}), we return a record type that has the union
of fields of the two arguments. We assume that the names of the first $k$ fields are the same and the
remaining fields have different names. To get a record with the right order of elements, we provide the
(\emph{order}) rule. The types of shared fields become common supertypes of their 
respective types (recursively). Fields that are present in only one record are marked as optional using
the following helper definition:
%
\begin{equation*}
\begin{array}{rcll}
 \addopt{\hat{\sigma}} &\narrow{=}& \hat{\sigma}~\kvd{option} \qquad&(\textnormal{non-nullable types})\\
 \addopt{\sigma} &\narrow{=}& \sigma &(\textnormal{otherwise})
\end{array}
\end{equation*}
%
Finding a common supertype of when one type is a union is more complicated. Our definition aims to 
use a flat structure (avoid nesting unions) and limit the number of cases. This is done by grouping
types that have a common supertype which is not a union type. For example, rather than inferring 
$\ident{int} + (\ident{bool} + \ident{decimal})$, our algorithm will find the common supertype of
$\ident{int}$ and $\ident{decimal}$ (which is \ident{decimal}) and it will produce just $\ident{decimal}+\ident{bool}$.

To identify which types have a common supertype which is not a union, we group the types by a 
\emph{tag}, which is defined as:
%
\begin{equation*}
\begin{array}{rcl}
 \tytag &\narrow{=}&  \ident{string} \lsep \ident{bool} \lsep \ident{number} \lsep \ident{union} \lsep \ident{collection} \\
        &\narrow{\lsep}& \ident{rec-anon} \lsep \ident{rec-named}\;~\nu 
\end{array}
\end{equation*}
%
The tag of a type is obtained using a function $\tytagof{(-)} : \sigma \rightarrow \tytag$:
%
\begin{equation*}
\begin{array}{rcl}
 \tytagof(\ident{string}) &\narrow{=}& \ident{string}\\
 \tytagof(\ident{bool}) &\narrow{=}& \ident{bool}\\
 \tytagof(\ident{decimal}) = \tytagof(\ident{float}) &\narrow{=}& \ident{number}\\
 \tytagof(\ident{int}) = \tytagof(\ident{bit}) &\narrow{=}& \ident{number}\\
 \tytagof([\sigma]) &\narrow{=}& \ident{collection}\\
\end{array}
\end{equation*}
\begin{equation*}
\begin{array}{rcl}
 \tytagof(\sigma + \ldots + \sigma) &\narrow{=}& \ident{union}\\
 \tytagof(\{ \nu_1 : \sigma_1, \; \ldots \; , \nu_n : \sigma_n \}) &\narrow{=}& \ident{rec-anon}\\
 \tytagof(\nu\; \{ \nu_1 : \sigma_1, \; \ldots \; , \nu_n : \sigma_n \}) &\narrow{=}& \ident{rec-named}~\nu \\
 \tytagof(\hat{\sigma}~\kvd{option}) &\narrow{=}& \tytagof(\hat{\sigma})
\end{array}
\end{equation*}
%
The function is undefined for the $\top$ and $\kvd{null}$ types, but this is not a 
problem because these types never need to be used as arguments in Figure~\ref{fig:subtyping-cst}.
The $\top$ type is always eliminated using the (\emph{top}) rule and the \kvd{null} type
is eliminated using either (\emph{null-1}) or (\emph{null-2}).

The handling of unions is specified using three rules. When adding a non-union type to a 
union (\emph{union-1}), the union may or may not contain a case with the same tag as the new type.
If it does, the new type is combined with the existing one, otherwise a new case is added.
When combining two unions (\emph{union-2}), we group the cases that have a shared tags.
Finally, the last rule (\emph{union-3}) covers the case when we are combining two non-union types.
As discussed earlier, union implicitly permits \kvd{null} values and so we use the following 
auxiliary function which makes nullable types non-nullable (when possible) to simplify the type:
%
\begin{equation*}
\begin{array}{rcll}
 \dropopt{\hat{\sigma}~\kvd{option}} &\narrow{=}& \hat{\sigma} &(\textnormal{option})\\
 \dropopt{\sigma} &\narrow{=}& \sigma &(\textnormal{otherwise})
\end{array}
\end{equation*}
%
The remaining rules are straightforward. For collections, we find the common supertype of 
its elements (\emph{list}); for compatible primitive types, we choose one of the two (\emph{prim})
and a common supertype with \kvd{null} is either the type itself (\emph{null-1}) or an option 
 (\emph{null-2}).

\paragraph{Properties.}
The common supertype relation finds a single supertype (among multiple candidates).
The following theorems specify that the found type is uniquely defined (up to reordering of 
fields in records and union cases) and is, indeed, a common supertype.
%
\begin{theorem}[Uniqueness]
\raggedright
If $\sigma_1 \triangledown \sigma_2 \vdash \sigma$ and $\sigma_1 \triangledown \sigma_2 \vdash \sigma'$
then it holds that $\sigma :> \sigma'$ and $\sigma' :> \sigma$.
\end{theorem}
\begin{proof}
The assumptions and shapes of the types to be unified in the rules in Figure~\ref{fig:subtyping-cst} 
are disjoint, with the exception of (\emph{order}), (\emph{sym}) and (\emph{refl}). These rules 
always produce types that are subtypes of each other. This property is preserved by rules that use the
common supertype relation recursively.
\end{proof}

\begin{theorem}[Common supertype]
\raggedright
If $\sigma_1 \triangledown \sigma_2 \vdash \sigma$ then it holds that $\sigma :> \sigma_1$ and $\sigma :> \sigma_2$.
\end{theorem}
\begin{proof}
By induction over the common supertype derivation $\vdash$.
\end{proof}

\noindent
We stated earlier that the common supertype relation minimises the use of union types 
(by preferring common numeric types or common record types when possible). This property
can be stated and proved formally:
%
\begin{theorem}
\label{thm:no-unions}
Given $\sigma_1, \sigma_2$ if there is s $\sigma$ such that $\sigma :> \sigma_1$ and 
$\sigma :> \sigma_2$ and $\sigma$ is not a union type, then $\sigma_1 \tsep \sigma_2 \vdash \sigma'$ 
and $\sigma'$ is not a union type.
\end{theorem}
\begin{proof}
The only rule that introduces an union type is (\emph{union-3}), which is used only when 
no other inference rule can be applied.
\end{proof}

% ==================================================================================================

\section{Formalising type providers}
\label{sec:formal}

In this section, we build the necessary theoretical framework for proving the relativised type safety 
property of F\# Data type providers in Section~\ref{sec:safety}. We start by discussing the runtime 
representation of structural values and runtime conversions (Section~\ref{sec:formal-convert}). Then 
we embed these into a formal model of an F\# subset that is necessary for discussing type providers 
(Section~\ref{sec:formal-ff}) and we describe how F\# Data type providers turn an inferred 
structural type into actual F\# code (Section~\ref{sec:formal-tp}).

% -------------------------------------------------------------------------------------------------

\subsection{Structural values and conversions}
\label{sec:formal-convert}

In the model presented here, we represent JSON, XML and CSV documents using the same 
\emph{structural value}\footnote{Here, we diverge from the actual implementation in F\# Data which 
uses a different implementation for each format (reusing existing libraries).}. Structural values are 
first-order and can be one of the primitive values (Boolean \kvd{true} or \kvd{false}, string $t$, 
integer $i$, decimal $d$ and floating-point number $f$), a missing value (\kvd{null}), a collection 
and a record (named or unnamed):
%
\begin{equation*}
\begin{array}{lcl}
 s  &\narrow{=}& i \lsep d \lsep f \lsep t \lsep \kvd{true} \lsep \kvd{false} \lsep \kvd{null} \\[0.1em]
    &\narrow{|}& [s_1; \ldots; s_n] \\[0.1em]
    &\narrow{|}& \{ \nu_1 \mapsto s_1, \ldots, \nu_n \mapsto s_n \}\\[0.1em]
    &\narrow{|}& \nu~\{ \nu_1 \mapsto s_1, \ldots, \nu_n \mapsto s_n \} \\
\end{array}
\end{equation*}

\noindent
The first few cases represent primitive values ($i$ for integers, $d$ for decimals, $f$ for floating
point numbers and $t$ for strings). A collection is written as a list of values in square brackets, 
separated by semicolons. A record can optionally start with its name $\nu$, followed by a sequence of 
field assignments $\nu_i \mapsto s_i$ written in curly brackets.

As suggested by subtyping, our system permits certain runtime conversions (\emph{e.g.}~a numerical 
value \num{1} can be treated as \kvd{true}). These are captured by the following operations:
%
\begin{equation*}
\begin{array}{lcl}
 op &\narrow{=}& \ident{asInt}(s) \lsep \ident{asDec}(s) \lsep \ident{asFloat}(s) \lsep \ident{asBool}(s)  \\
    &\narrow{|}& \ident{asStr}(s) \lsep \ident{getChildren}(s) \lsep \ident{getField}(\nu?, \nu, s)\\
    &\narrow{|}& \ident{isNull}(s) \lsep \ident{isTag}(\ident{tag}, s)
\end{array}
\end{equation*}
%
The reduction rules for these operations are defined in Figure~\ref{fig:op-conversions}. It is important
to note that the relativised property we prove about F\# Data specifies \emph{sufficient}, but
not \emph{necessary} conditions. That is, a type provided based on a sample is guaranteed to work 
correctly on certain inputs, but may also handle several additional inputs.

The conversion functions follow this approach. They perform conversions required by the subtyping relation
(\emph{e.g.}~integer is converted to decimal or a float) but also attempt additional conversions.

\begin{itemize}
\item \textbf{Widening operations.} The widening operations are required by the subtyping. These include
  converting to a wider numerical type (integer to decimal and float, decimal to float, converting 0 and 
  1 to Boolean).

\item \textbf{Conversions.} Additional conversions are not strictly required, but prove useful in practice.
  These include converting to a smaller numerical type (such as float to integer) written as $i=\lfloor f\rfloor$,
  which is not defined in case of arithmetic overflow; conversion to and from string values written as
  ``$t$~represents~$i$'', which parses or formats a non-string value and also treating of decimals 
  0.0 and 1.0 as Booleans.

\item \textbf{Structural operations.} Aside from primitives, the \ident{getChildren} operation returns treats
  \kvd{null} as an empty collection. The \ident{getField} function is defined separately for named and unnamed
  records (we write $\bot$ for a name of an unnamed record). Accessing a field of a named record checks that 
  the expected record name matches the actual.

\item \textbf{Helper functions.} The semantics also defines two helper functions. The \ident{isNull} function
  is used to check whether value is \kvd{null} and \ident{isTag} is used to test whether a value can be treated 
  as a value of type associated with the specified tag (as defined in Section~\ref{sec:inference-commonsuper}).
  The last line defines a ``catch all'' pattern, which returns \kvd{false} for all remaining cases.
\end{itemize}

% -------------------------------------------------------------------------------------------------

\begin{figure}
\begin{equation*}
\begin{array}{rl}
 \textnormal{\footnotesize{(member)}}&
 \hspace{-0.4em}
 \inference
 { \kvd{type}~C(\overline{x:\tau})=\ldots \kvd{member}~N_i : \tau_i = e_i \ldots }
 { (\kvd{new}~C(\overline{v})).N_i \reduce e_i[\overline{x} \leftarrow \overline{v}] }\\
 \\
 \textnormal{\footnotesize{(ctx)}}&
 \hspace{-0.4em}
  E[e] \reduce E[e'] \qquad\qquad(\textnormal{when}~e \reduce e')\\
 \\
 \textnormal{\footnotesize{(cond1)}}&
 \hspace{-0.4em}
 \kvd{if}~\kvd{true}~\kvd{then}~e_1~\kvd{else}~e_2 ~\reduce~ e_1 \\
 \\
 \textnormal{\footnotesize{(cond2)}}&
 \hspace{-0.4em}
 \kvd{if}~\kvd{false}~\kvd{then}~e_1~\kvd{else}~e_2 ~\reduce~ e_2 \\
 \\
 \textnormal{\footnotesize{(match1)}}&
 \hspace{-1em}
 \begin{array}{l}
  \kvd{match}~\ident{None}~\kvd{with} \\
  \ident{Some}(x) \rightarrow e_1 \,|\, \ident{None} \rightarrow e_2
 \end{array} \hspace{-0.5em} ~\reduce~ e_2 \\
 \\
 \textnormal{\footnotesize{(match2)}}&
 \hspace{-1em}
 \begin{array}{l}
    \kvd{match}~\ident{Some}(v)~\kvd{with} \\
    \ident{Some}(x) \rightarrow e_1 \,|\, \ident{None} \rightarrow e_2
 \end{array} \hspace{-0.5em} ~\reduce~ e_1[x\leftarrow v]\\
 \\
 \textnormal{\footnotesize{(match3)}}&
 \hspace{-1em}
 \begin{array}{l}
  \kvd{match}~[v_1;\ldots;v_n]~\kvd{with} \\[0em]
  [x_1;\ldots;x_m ] \rightarrow e_1 \,|\, \_ \rightarrow e_2
 \end{array} \hspace{-0.5em} ~\reduce~ e_2\quad(m\neq n) \\
 \\
 \textnormal{\footnotesize{(match4)}}&
 \hspace{-1em}
 \begin{array}{l}
  \kvd{match}~[v_1;\ldots;v_n]~\kvd{with} \\[0em]
  [x_1;\ldots;x_n ] \rightarrow e_1 \,|\, \_ \rightarrow e_2
 \end{array} \hspace{-0.5em} ~\reduce~ e_1[\overline{x}\leftarrow\overline{v}] \\
 \\
 \textnormal{\footnotesize{(map)}}&
 \hspace{-0.4em}
 \ident{List.map}~(\lambda x.e)~[v_1; \ldots] ~\reduce~ [e[x\leftarrow v_1]; \ldots] \\
\end{array}
\end{equation*}

\caption{Featherweight F\# -- Remaining reduction rules}
\label{fig:ff-reduction}
\vspace{-1em}
\end{figure}

% -------------------------------------------------------------------------------------------------

\subsection{Featherweight F\#}
\label{sec:formal-ff}

The semantics fragment introduced in the previous section discusses values and operations that are used by
F\# Data. This section adds a minimal subset of the standard F\# language.

We focus on the object system (which is what F\# type providers produce) and so the following combines 
standard ML \cite{sml} features with Featherweight Java \cite{fwjava} classes. However, we only need classes with properties and
without inheritance. A class has a single implicit constructor and the declaration closes over 
constructor parameters. To avoid including all of ML, we only select constructs for working with options and lists
that we need later.
%
\begin{equation*}
\begin{array}{rcl}
 \tau &\narrow{=}& \ident{int} \lsep \ident{decimal} \lsep \ident{float} \lsep \ident{bool} \lsep \ident{string} \\[0.0em]
      &\narrow{|}& C \lsep \ident{StructVal} \lsep \ident{list}\langl\tau\rangl \lsep \ident{option}\langl\tau\rangl \\[0.6em]
 L &\narrow{=}& \kvd{type}~C(\overline{x:\tau}) = \overline{M} \\[0.0em]
 M &\narrow{=}& \kvd{member}~N:\tau=e \\[0.6em]
 v &\narrow{=}& s \lsep \ident{None} \lsep \ident{Some}(v) \lsep \kvd{new}~C(\overline{v}) \lsep [v_1; \ldots; v_n] \\[0.0em]
 e &\narrow{=}& op \lsep e.N \lsep \kvd{new}~C(\overline{e}) \lsep {\kvd{if}~e_1~\kvd{then}~e_2~\kvd{else}~e_3}\\
   &\narrow{|}& \ident{Some}(e)\lsep\kvd{match}~e~\kvd{with}~\ident{Some}(x) \rightarrow e_1 \,|\, \ident{None} \rightarrow e_2 \\
   &\narrow{|}& \kvd{match}~e~\kvd{with}~[x_1; \ldots; x_n] \rightarrow e_1 \,|\, \_ \rightarrow e_2 \\
   &\narrow{|}& [e_1;\ldots;e_n]\lsep \ident{List.map}~(\lambda x\rightarrow e_1)~e_2
\end{array}
\end{equation*}

\noindent
In the above definition, $\tau$ denotes types including a type of all structural values $s$ called
\ident{StructVal}. A class definition $L$ consists of a constructor and zero or more members. Values $v$ include 
previously defined structural values $s$ and values for the option and list type; finally expressions $e$ include 
previously defined operations $op$, class construction, member access, conditionals and expressions for working 
with option values and lists. We include \ident{List.map} as a special construct to avoid making the language too complex.

Next, we define the reduction relation and (a fragment of) type checking for Featherweight F\#.
The language presented here is intentionally incomplete. We only define parts needed to prove
the relativized safety property in Section~\ref{sec:safety}.

\paragraph{Reduction.} The reduction relation is of the form $e \reduce e'$. We also write 
$e \reduce^{*} e'$ to denote the reflexive and transitive closure of $\reduce$. The reduction rules
for primitive functions $op$ were discussed earlier in Figure~\ref{fig:op-conversions}. The rules 
for other expressions defined above are provided in Figure~\ref{fig:ff-reduction}.

The (\emph{ctx}) rule performs a reduction inside a sub-expression specified by an evaluation context.
This models the eager evaluation order of F\#. An evaluation context $E$ is defined as:

\vspace{0.5em}
\noindent
\begin{equation*}
\begin{array}{rcl}
 E &\narrow{=}& f(E) \lsep E.N \lsep \kvd{new}~C(\overline{v}, E, \overline{e}) \lsep \kvd{if}~E~\kvd{then}~e_1~\kvd{else}~e_2 \\[0.1em]
   &\narrow{|}& \ident{Some}(E) \lsep \kvd{match}~E~\kvd{with}~\ident{Some}(x) \rightarrow e_1 \,|\, \ident{None} \rightarrow e_2 \\[0.1em]
   &\narrow{|}& [\overline{v};E;\overline{e}] \lsep \kvd{match}~E~\kvd{with}~[x_1; \ldots; x_n] \rightarrow e_1 \,|\, \_ \rightarrow e_2 \\[0.1em]
   &\narrow{|}& \ident{map}~(\lambda x\rightarrow e)~E \lsep [\cdot] 
 \\[0.6em]
 f &\narrow{\in}& \{\; \ident{asInt}, \ident{asDec}, \ident{asFloat}, \ident{asBool}, \ident{asStr}, \\[0.1em]
              && ~~\; \ident{getField}, \ident{getChildren}, \ident{isNull}, \ident{isTag} \;\}
\end{array} 
\end{equation*}

\noindent
The evaluation first reduces arguments of functions and the evaluation proceeds from left to right 
as denoted by $\overline{v}, E, \overline{e}$ in constructor arguments or $\overline{v};E;\overline{e}$
in list initialization.

We write $e[\overline{x} \leftarrow \overline{v}]$ for the result of replacing variables $\overline{x}$ by
values $\overline{v}$ in expression. The (\emph{member}) rule reduces a member access using a class 
definition in the assumption to obtain the body of the expression. Finally, the remaining six rules
provide standard reductions for conditionals and pattern matching.

The language is sufficient for our purpose, but simple. It does not provide recursion
and so all expressions reduce to a value in a finite number of steps or get stuck due to an error
condition. An error condition can be a wrong argument passed to conditional, pattern matching or
one of the conversion functions from Figure~\ref{fig:op-conversions}.

% -------------------------------------------------------------------------------------------------

\begin{figure}
\begin{equation*}
\qquad
\inference
  {~}
  {L; \Gamma \vdash s : \ident{StructVal}}
\qquad  
\inference
  {~}
  {L; \Gamma \vdash n : \ident{int}}
\end{equation*}

\begin{equation*}
\inference
  {L; \Gamma \vdash e : C & \kvd{type}~C(\overline{x:\tau}) = ..\;\kvd{member}~N_i : \tau_i = e_i\;.. \in L}
  {L; \Gamma \vdash e.N_i:\tau_i}
\end{equation*}

\begin{equation*}
\inference
  {L; \Gamma \vdash e_i : \tau_i & \kvd{type}~C(x_1:\tau_1, \ldots, x_n:\tau_n) = \ldots \in L}
  {L; \Gamma \vdash \kvd{new}~C(e_1, \ldots, e_n):C}
\end{equation*}

\caption{Featherweight F\# -- Fragment of type checking}
\label{fig:ff-typecheck}
\end{figure}

% -------------------------------------------------------------------------------------------------

\paragraph{Type checking.} 
The type checking rules in Figure~\ref{fig:ff-typecheck} are written using a judgement
$L; \Gamma \vdash e : \tau$ where the context also contains a set of class declarations $L$.
The rules demonstrate the key difference between our language and standard ML or Featherweight Java:
%
\begin{itemize}[noitemsep]
\item[--] All structural values $s$ have a type \ident{StructVal}. Some of those have other types
  (primitive integers, decimals, floats, strings and Booleans as illustrated by the rule for $n$).
  For other values, \ident{StructVal} is the only type -- this includes records and \kvd{null}.
\item[--] A list containing other structural values $[s_1; \ldots; s_n]$ has a type \ident{StructVal}.  
  but can also have the $\ident{list}\langl\tau\rangl$ type. Conversely, lists that contain
  non-structural values like objects or options are not of type \ident{StructVal}.
\item[--] Primitive operations $op$ are treated as functions, accepting \ident{StructVal} and producing
  an appropriate type as the result.
\item[--] The two rules for checking class construction and member access are similar to corresponding
  rules of Featherweight Java.  
\end{itemize}
%
An important part of Featherweight Java that is omitted here is the checking of type declarations
(ensuring the bodies of members are well-typed). We omit this, because it is not needed for the
relativized safety theorem, which considers only classes generated by our type provider (which 
are correct by construction).

% -------------------------------------------------------------------------------------------------

\begin{figure*}
\begin{multicols}{2}
% Primitive structural types become corresponding F# types and conversion is inserted
\noindent
\begin{equation*}
\begin{array}{l}
 \sem{\sigma_p}_e = \tau_p,op(e),\emptyset\quad\textnormal{where} \\[0.4em]
\quad\sigma_p, \tau_p, op\in  \{~ (\ident{bit}, \ident{bool}, \ident{asBool}),  (\ident{bool}, \ident{bool}, \ident{asBool})\\[0.2em]
\hspace{2.9em} (\ident{int}, \ident{int}, \ident{asInt}), (\ident{decimal},\ident{decimal},\ident{asDec}),\\[0.2em]
\hspace{2.9em} (\ident{float},\ident{float},\ident{asFloat}), (\ident{string},\ident{string},\ident{asStr}) ~\}\\[0.6em]
\end{array}
\end{equation*}
%
% Records become classes
\begin{equation*}
\begin{array}{l}
 \sem{\nu?\; \{ \nu_1 : \sigma_1, \ldots, \nu_n : \sigma_n \}}_e = \\[0.1em]
 \quad C, \kvd{new}~C(e), L_1\cup\ldots\cup L_n\cup\{ L \}\quad\textnormal{where}\\[0.6em]
 \qquad \;\;C~\textnormal{is a fresh class name} \\[0.1em]
 \qquad \;\,\,L = \kvd{type}~C(v:\ident{StructVal})~=~M_1\ldots M_n  \\[0.1em]
 \qquad M_i = \kvd{member}~\nu_i:\tau_i=e_i\\[0.1em]
 \qquad \tau_i, e_i, L_i = \sem{\sigma_i}_{e'},\;\; e'=\ident{getField}(\nu?, \nu_i, e)\\[0.6em]
\end{array}
\end{equation*}
%
% Collections
\begin{equation*}
\begin{array}{l}
 \sem{\,[\sigma]\,}_e = \ident{list}\langl\tau\rangl, \ident{List.map}~(\lambda x\rightarrow e')~(\ident{getChildren}(e)), L\\[0.4em]
 \quad \textnormal{where}~~\tau, e', L = \sem{\hat{\sigma}}_x
\end{array}
\end{equation*}

% Option values
\noindent
\begin{equation*}
\begin{array}{l}
 \sem{\hat{\sigma}\;\,\kvd{option}}_e = \\[0.2em]
 \hspace{1.25em} \ident{option}\langl\tau\rangl, \kvd{if}~\ident{isNull}~e~\kvd{then}~\ident{None}~\kvd{else}~\ident{Some}(e'), L\\[0.2em] 
 \hspace{1.25em} \textnormal{where}~\tau, e', L = \sem{\hat{\sigma}}_e
\end{array}
\end{equation*}
%
% Sum type
\begin{equation*}
\begin{array}{l}
 \sem{\sigma_1 + \ldots + \sigma_n}_e = \\[0.1em]
 \quad C, \kvd{new}~C(e), L_1\cup\ldots\cup L_n\cup\{ L \}\quad\textnormal{where}\\[0.6em]
 \qquad \;\;C~\textnormal{is a fresh class name} \\[0.1em]
 \qquad \;\,\,L = \kvd{type}~C(v:\ident{StructVal})~=~M_1\ldots M_n \\[0.1em]
 \qquad M_i = \kvd{member}~\nu_i:\ident{option}\langl\tau_i\rangl=\\[0.1em]
 \hspace{5.8em}  \kvd{if}~\ident{isNull}(v)~\kvd{then}~\ident{None}~\kvd{else} \\
 \hspace{5.8em}  \kvd{if}~\ident{isTag}(t_i, v)~\kvd{then}~
     \ident{Some}(e_i)~\kvd{else}~\ident{None} \\[0.1em]
 \qquad \tau_i, e_i, L_i = \sem{\sigma_i}_e,\;\; t_i = \tytagof{(\sigma_i)},\;\; \nu_i=\nameoftag{(t_i)}
\end{array}
\end{equation*}
%
% Anything else
\begin{equation*}
\begin{array}{l}
 \sem{\top}_v = \sem{\kvd{null}}_v = \ident{StructVal}, v, \emptyset
\end{array}
\end{equation*}
\end{multicols}

\vspace{-0.5em}
\noindent
\begin{equation*}
\qquad
\qquad
\begin{array}{rcl}
 \nameoftag(\ident{string}) &\narrow{=}& \ident{String} \\
 \nameoftag(\ident{bool}) &\narrow{=}& \ident{Boolean} \\
\end{array}
\qquad
\begin{array}{rcl}
 \nameoftag(\ident{number}) &\narrow{=}& \ident{Number} \\
 \nameoftag(\ident{collection}) &\narrow{=}& \ident{List} \\
\end{array}
\qquad
\begin{array}{rcl}
 \nameoftag(\ident{rec-anon}) &\narrow{=}& \ident{Record} \\
 \nameoftag(\ident{rec-named}\;\nu) &\narrow{=}& \nu \\
\end{array}
\end{equation*}

\caption{Type provider -- generation of featherweight F\# types from inferred structural types}
\label{fig:tp-generation}
\vspace{-0.5em}
\end{figure*}

% -------------------------------------------------------------------------------------------------

\subsection{Type providers}
\label{sec:formal-tp}

So far, we defined the type inference algorithm for structured data formats which produces a structural
type $\sigma$ based on one or more sample documents (Section~\ref{sec:inference}). In Section~\ref{sec:formal-ff},
we defined a simplified model of F\# language and evaluation, which also included conversions that
model the F\# Data library runtime (Section~\ref{sec:formal-tp}). In this section, we define how
the type providers work, which links the two aspects discussed so far.

The type providers for XML, CSV and JSON all work in the same way. They take sample documents and infer a common
supertype $\sigma$ from the samples. From $\sigma$, they then generate F\# types that are then exposed to the
programmer\footnote{The actual implementation provides \emph{erased types} as described in \cite{fsharp-typeprov}. 
Here, we treat the code as actually generated. This is an acceptable simplification, because F\# Data type providers do not 
rely on laziness that is available through erased types.}.

\paragraph{Type provider mapping.}
When generating types, the type provider produces an F\# type $\tau$, expression that wraps a structural 
document value (of type \ident{StructVal}) as a value of type $\tau$ and a collection of class definitions. 
We express it using the following mapping:
%
\begin{equation*}
\sem{-}_e : \sigma \rightarrow \tau \times e' \times L
\end{equation*}
%
The semantics is parameterized by an expression $e$, which represents code to obtain a structural value that is
being wrapped. Then, given an inferred structural type $\sigma$, the type provider produces an F\# type $\tau$,
an expression $e'$ which constructs a value of type $\tau$ using $e$ and also a set of class definitions $L$.

Figure~\ref{fig:tp-generation} shows the rules that define $\sem{-}_e$. Primitive types are all
handled by a single rule. For a given structural type, it returns the corresponding F\# type -- the types are
mapped directly with the exception of \ident{bit}, which becomes an F\# \ident{bool}. The generated code calls
an appropriate conversion function from Figure~\ref{fig:op-conversions} on the input.

Handling of records is more interesting. We generate a new class $C$ that takes a structural value as constructor
parameter. For each record field, we generate a new member with the same name as the field\footnote{The actual
F\# Data implementation also capitalizes the names.} The body of the member calls \ident{getField} and then
passes this expression to $\sem{\sigma_i}$ which adds additional wrapping that maps the field (structural value
of type $\sigma_i$) into an F\# type $\tau_i$. The returned expression creates a new instance of $C$ and
the mapping returns the class $C$ together with all recursively generated definitions.

A collection type becomes an F\# $\ident{list}\langl\tau\rangl$. The returned expression calls \ident{getChildren}
(which turns \kvd{null} values into empty lists) and then uses \ident{List.map} to convert all nested values to 
an F\# type $\tau$. The handling of option type is similar -- it checks if the original value is \kvd{null} and
if no, it wraps the recursively generated conversion expression $e'$ in the \ident{Some} constructor.

As discussed earlier, union types are also generated as classes with properties. Given a union type $\sigma_1 + \ldots + \sigma_n$,
we get corresponding F\# types $\tau_i$ and generate $n$ members of type $\ident{option}\langl \tau_i\rangl$.
Each member return \ident{Some} when the value is not \kvd{null} and has the right structure (checked by 
\ident{isTag}). The type inference algorithm also guarantees that there is only one case for each type tag 
(Section~\ref{sec:inference-commonsuper}) and there are no nested union types. Thus, checking for a tag is
sufficient and we can also use the tag to identify the name of the generated member (using the
\ident{nameof} function).

\paragraph{Example 1.}
To illustrate how the type provision mechanism works, we now consider two simple examples. First, assume 
that the inferred type is a record with two fields (one optional) such as 
$\ident{Person}~\{~\ident{Age}:\ident{int}~\kvd{option}, \ident{Name}:\ident{string}~ \}$. Applying the
rules from Figure~\ref{fig:tp-generation} produces the following class:
%
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{Person}(\ident{v}:\ident{StructVal})~= \\[0.1em]
 \quad \kvd{member}~\ident{Age}~:~\ident{option}\langl\ident{int}\rangl~= \\[0.1em]
 \qquad \kvd{if}~\ident{isNull}~(\ident{getField}(\ident{Person},\ident{Age}, \ident{v}))~\kvd{then}~\ident{None} \\[0.1em]
 \qquad \kvd{else}~\ident{Some}(\ident{asInt}(\ident{getField}(\ident{Person},\ident{Age}, \ident{v}))) \\[0.1em]
 \quad \kvd{member}~\ident{Name}~:~\ident{string}~= \\[0.1em]
 \qquad \ident{asStr}(\ident{getField}(\ident{Person},\ident{Name}, \ident{v}))
\end{array}
\end{equation*}
%
The body of the \ident{Age} member is produced by the case for optional types applied to an expression
$\ident{getField}(\ident{Person},\ident{Age},v)$. If the returned field is not \kvd{null}, then the member
calls \ident{asInt} (produced by the case for primitive types) and wraps the result in the \ident{Some}
constructor. Note that \ident{getField} is defined even when the field does not exist, but returns \ident{null}.
This lets us treat missing fields as optional fields. The \ident{Name} member is similar, but does not 
perform any checks. 
For completeness, a type corresponding to the record is \ident{Person} and given a 
structural value $s$, we create a \ident{Person} value by calling $\kvd{new}~\ident{Person}(s)$.

\paragraph{Example 2.} The second example illustrates the remaining parts of the type provision,
including collections and union types. Reusing the \ident{Person} type from the previous example,
consider a collection $[\ident{Person}+\ident{string}]$, which contains a mix of \ident{Person} and string values.

\noindent
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{PersonOrString}(\ident{v}:\ident{StructVal})~= \\[0.1em]
 \quad \kvd{member}~\ident{Person}~:~\ident{option}\langl\ident{Person}\rangl~= \\[0.1em]
 \qquad \kvd{if}~\ident{isNull}(v)~\kvd{then}~\ident{None}~\kvd{else} \\[0.1em]
 \qquad \kvd{if}~\ident{isTag}(\ident{rec-named Person}, v)~\kvd{then}~\\[0.1em]
 \qquad\quad \ident{Some}(\kvd{new}~\ident{Person}(v))~\kvd{else}~\ident{None} \\[0.1em]
 \quad \kvd{member}~\ident{String}~:~\ident{option}\langl\ident{string}\rangl~= \\[0.1em]
 \qquad \kvd{if}~\ident{isNull}(v)~\kvd{then}~\ident{None}~\kvd{else} \\[0.1em]
 \qquad \kvd{if}~\ident{isTag}(\ident{string}, v)~\kvd{then}~\\[0.1em]
 \qquad\quad \ident{Some}(\ident{asStr}(v))~\kvd{else}~\ident{None}
\end{array}
\end{equation*}

\noindent
The type provider generates the above type and the collection type is mapped to an F\# type 
$\ident{list}\langl\ident{PersonOrString}\rangl$. Given a structural document value $s$, the code
to obtain the wrapped F\# value is:
%
\begin{equation*}
\ident{List.map}~(\lambda x\rightarrow\kvd{new}~\ident{PersonOrString}(x))~(\ident{getChildren}(s))
\end{equation*}
%
The \ident{PersonOrString} type contains one member for each of the union case. In the body, they
check that the value is not \kvd{null} and that it has the right structure (using the \ident{isTag}
function). This checks that the value is a record named \ident{Person} or a string, respectively.
If the conditions are satisfied, the value is converted to the F\# type corresponding to the case
and wrapped in \ident{Some}.

% --------------------------------------------------------------------------------------------------

\subsection{Inferring types from values}
\label{sec:formal-inferval}

Before concluding the section on formalizing type providers, there is one more missing piece. The
common supertype algorithm in Section~\ref{sec:inference} discusses the core of the type inference for
structural values, but we have not yet clarified how exactly it is used. We address this in the present
section.

Given a JSON, XML or CSV document, the F\# Data implementation constructs a structural value $s$
from the sample (this is straightforward, but Section~\ref{sec:impl-parsing} discusses some interesting 
aspects). The following defines a mapping $\semalt{-}$ which turns a sample value $s$ into a structural
type $\sigma$:
%
\begin{equation*}
\begin{array}{rclcrcl}
 \semalt{0} &\narrow{=}& \ident{bit} &\qquad& \semalt{1} &\narrow{=}& \ident{bit} \\
 \semalt{i} &\narrow{=}& \ident{int} && \semalt{d} &\narrow{=}& \ident{decimal} \\
 \semalt{f} &\narrow{=}& \ident{float} && \semalt{t} &\narrow{=}& \ident{string} \\
 \semalt{b} &\narrow{=}& \ident{bool} && \semalt{\kvd{null}}  &\narrow{=}& \kvd{null} \\[0em]
\end{array}
\end{equation*}
\noindent
\vspace{-0.2em}
\begin{equation*}
\begin{array}{l}
 \semalt{[s_1; \ldots; s_n]} = [\semalt{s_1, \ldots, s_n}]
 \\[0.5em]
 \semalt{\nu?~\{ \nu_1 \mapsto s_1, \ldots, \nu_n \mapsto s_n \}} =\\[0.1em]
 \qquad\nu?~\{ \nu_1:\semalt{s_1}, \ldots, \nu_n :\semalt{s_n} \}
 \\[0.5em]
 \semalt{s_1, \ldots, s_n} = \sigma_n \quad\textnormal{where}\\[0.1em]
 \qquad\sigma_0 = \top,~ \forall i\in \{ 1.. n \}.~ \sigma_{i-1} \triangledown \semalt{s_i} \vdash \sigma_i
\end{array}
\end{equation*}
%
Primitive values are mapped to their corresponding types, with the exception of $0$ and $1$, which
are inferred as \ident{bit}. For records, we return a type with field types inferred based on 
individual values.

The interesting part is inferring type based on multiple samples. We overload
the notation and write $\semalt{s_1, \ldots, s_n}$ for a type inferred from multiple samples.
This uses the common supertype relation to find a common type for all values (starting with $\top$).
This operation is used at the top-level (when calling type provider with multiple samples)
and also when inferring the type of a collection.

% ==================================================================================================

\section{Relativized type safety}
\label{sec:safety}

Informally, the safety property of F\# Data type providers states that, given representative sample
documents, any code that can be written using the provided types is guaranteed to work. We call this 
\emph{relativized safety}, because we cannot avoid \emph{all} errors. In particular, the user can 
always use the type provider with an input that has a different structure than any of the samples -- and in 
this case, it is expected that the code will fail at runtime (throw an exception in the actual 
implementation or by getting stuck in our model).

The key question is, what is a representative sample? Given a set of sample documents, 
the provided type is guaranteed to work if the inferred type of the input is a subtype of any of the
sample documents. Going back to Section~\ref{sec:inference-subtyping}, this means that:
%
\begin{itemize}[noitemsep]
\item[--] Input can contain smaller numerical values (for example, if a sample contains float, the input can contain an integer).
\item[--] Records in the actual input can have additional fields
\item[--] Records in the actual input can have fewer fields, provided that the type of the fields is marked as optional in the sample
\item[--] Union types in the input can have both fewer or more cases
\item[--] When we have a union type in the sample, the actual input can also contain just values of one of the union cases
\end{itemize}
%
The following lemma states that the provided code (generated in Figure~\ref{fig:tp-generation})
works correctly on inputs $s'$ that is a subtype of the sample $s$. More formally, we require that the 
provided expression (using $s'$ as the input) can be reduced to a value and, if it is a class,
all its members can also be reduced to values.

\begin{lemma}[Correctness of provided types]
\label{thm:tp-correctness}
Given a sample value $s$ and an input value $s'$ such that $\semalt{s} :> \semalt{s'}$
and provided type, expression and class declarations $\tau, e, L = \sem{\semalt{s}}_{s'}$, 
then $e \reduce^{*} v$ and if $\tau$ is a class ($\tau=C$) then for all members $N_i$ of the 
class $C$, it holds that $e.N_i \reduce^{*} v$.
\end{lemma}
\begin{proof}
By induction over the structure of $\sem{-}_e$. For primitives, the conversion functions accept all subtypes.
For other cases, analyse the the provided code to see that it can work on all subtypes (\emph{e.g.}~\ident{getChildren}
works on \kvd{null} values, \ident{getField} returns \kvd{null} when a field is missing and, when \ident{isTag} returns
\ident{true}, then a value has the structure required by the corresponding case).
\end{proof}

\noindent
Now that we know that the provided types are correct with respect to the subtyping relation, we can
look at the main theorem of the paper. It states that, for any input (which is a subtype of any of the
samples) and any expression $e$, a well-typed program that uses the provided types does not ``go wrong''.

Using the standard approach to syntactic type safety  \cite{syntactic}, we prove the type preservation 
(reduction does not change type) and progress (an expression that is not a value can be reduced).

\begin{theorem}[Relativized safety]
\label{thm:safety}
Assume $s_1, \ldots, s_n$ are samples, $\sigma=\semalt{s_1, \ldots, s_n}$ is an inferred
type and $\tau,e,L = \sem{\sigma}_x$ are a type, expression and class definitions generated by a 
type provider.

Then for all new inputs $s'$ such that $\exists i.(\semalt{s_i} :> \semalt{s'})$, let $e_s=e[x\leftarrow s']$
be an expression (of type $\tau$) that wraps the input in a provided type. Then, for any expression $e_c$
(user code) such that $\emptyset; y:\tau \vdash e_c:\tau'$ and that does not contain any structural values
as sub-expressions, it is the case that $e_c[y\leftarrow e_s] \reduce^{*} v$ for some value $v$ and
also $\emptyset; \vdash v : \tau$.
\end{theorem}
\begin{proof}
We discuss the two parts of the proof separately as type preservation (Lemma~\ref{thm:rs-preservation})
and progress (Lemma~\ref{thm:rs-progress}).
\end{proof}

\begin{lemma}[Preservation]
\label{thm:rs-preservation}
Given the class definitions $L$ generated by a type provider as specified in
the assumptions of Theorem~\ref{thm:safety}, then if $\Gamma \vdash e : \tau$ and 
$e \reduce^{*} e'$ then $\Gamma \vdash e' : \tau$.
\end{lemma}
\begin{proof}
By induction over the reduction $\reduce$. The cases for the ML subset of Featherweight F\# 
are standard. For (\emph{member}), we check that code generated by type providers
in Figure~\ref{fig:tp-generation} is well-typed.
\end{proof}

\noindent
The progress lemma states that evaluation of a well-typed program does not reach an undefined state. 
This is not a problem for the ML subset and object-oriented subset of the calculus. The problematic
part are the conversion functions (Figure~\ref{fig:op-conversions}). Given a structural value 
(which has a type \ident{StructVal} in our language), the reduction can get stuck if the value does
not have a structure required by a specific conversion function used.

The Lemma~\ref{thm:tp-correctness} guarantees that this does not happen. In Theorem~\ref{thm:safety},
we carefully state that we only consider expressions $e_c$ which ``[do] not contain any structural 
values as sub-expressions''. This makes sure that the only code working with structural values is the
code generated by the type provider.

\begin{lemma}[Progress]
\label{thm:rs-progress}
Given the assumptions and definitions from Theorem~\ref{thm:safety}, it is the case that
$e_c[y\leftarrow e_s] \reduce^{*} v$.
\end{lemma}
\begin{proof}
By induction over the typing derivation of $L; \emptyset \vdash e_c[y\leftarrow e_s] : \tau'$. 
The cases for the ML subset are standard. For member access, we rely on Lemma~\ref{thm:tp-correctness}.
\end{proof}

% -------------------------------------------------------------------------------------------------

\subsection{Discussion}
\label{sec:safety-discuss}

The \emph{relativized safety} property does not guarantee the same amount of safety as standard type safety
for programming languages without type providers. However, it reflects the reality of programming with external
data sources that is increasingly important in the age of web \cite{age-of-web}. So, type providers do not
reduce the safety -- they simply make the existing issues visible.

The actual implementation in F\# Data throws an exception for invalid inputs. In contrast, the calculus
presented here simply gets stuck. We could extend the calculus with exceptions, but that would obscure the 
purpose -- precisely specifying when the code using type providers does not ``go wrong.''

As mentioned earlier, \emph{relativized safety} specifies a sufficient, but not a necessary condition.
This is also reflected in the conversion functions which allow additional conversions not required by 
the subtyping relation. In practice, the additional flexibility proves useful (as minor variations in 
input formats are common). However, an interesting stricter alternative would be to check that an input 
has the required type when it is loaded -- and throw an exception immediately when loading data rather than 
on member access.

Finally, the formal model presented here ignores two aspects of the real type provider mechanisms. As discussed
here, we do not use the \emph{type erasure} mechanism, but treat type providers as if they were actually 
generating code. For XML, CSV and JSON, both models can be used. However, one interesting aspect of erasing 
type providers is that the types can be generated lazily -- only classes that are actually used in the type
checked code are generated. Extending our formalism to capture this aspect is an interesting future work
relevant to other type providers -- including the World Bank type provider that is also available
in the F\# Data library.

% ==================================================================================================
%
%   ###                                                                                     
%    #  #    # #####  #      ###### #    # ###### #    # #####   ##   ##### #  ####  #    # 
%    #  ##  ## #    # #      #      ##  ## #      ##   #   #    #  #    #   # #    # ##   # 
%    #  # ## # #    # #      #####  # ## # #####  # #  #   #   #    #   #   # #    # # #  # 
%    #  #    # #####  #      #      #    # #      #  # #   #   ######   #   # #    # #  # # 
%    #  #    # #      #      #      #    # #      #   ##   #   #    #   #   # #    # #   ## 
%   ### #    # #      ###### ###### #    # ###### #    #   #   #    #   #   #  ####  #    # 
%
% ==================================================================================================

\section{Implementation}
\label{sec:impl}

The theoretical model in Section~\ref{sec:formal} presents the core ideas behind the type 
providers for structured data formats. However, an important part of the success of the F\# Data
library is also its pragmatic approach, which makes it suitable for use with real-world data.

The formal model discusses some of the pragmatic choices, such as the preference for records over 
unions and the ability to treat $0$ and $1$ as both Booleans and numbers. In this section, we briefly
discuss some of the remaining practical concerns, starting with the handling of collections.

% -------------------------------------------------------------------------------------------------

\subsection{Parsing structured data}
\label{sec:impl-parsing}

In this paper, we treat the XML, JSON and CSV formats uniformly as \emph{structural values}.
The definition of structural values is, indeed, rich enough to capture all three formats. The
structure is similar to JSON (it has primitive values, records and collections). We also added
\emph{named} records, which are needed for XML.
%
\begin{itemize}
\item When reading JSON, we directly create a corresponding structural value (with all records unnamed).
  Optionally, we also read numerical values (and dates) stored in strings.
\item When reading CSV, we read each row as an unnamed record and return them in a collection.
  Optionally, the read values are parsed as numbers or Booleans and missing values become \kvd{null}.
\item When reading XML, we create a named record for each node. Attributes become record fields and
  body becomes a special field named $\bullet$
\end{itemize}
%
The reading of XML documents is perhaps the most interesting. To demonstrate how the body is treated,
consider the following:
%
{\small{
\begin{verbatim}
  <root id="1">
    <item>Hello!</item>
  </root>    
\end{verbatim}
}}
%
\noindent
This XML becomes a record named \ident{root} with fields \ident{id} and $\bullet$. The nested element
contains only the $\bullet$ field containing the inner text:
%
\begin{equation*}
\ident{root}~\{ \ident{id} \mapsto 1, \bullet \mapsto [ \ident{item}~\{ \bullet \mapsto \str{Hello!} \}] \}
\end{equation*}
%
When generating types for types inferred from XML, we also include a special case to remove the $\bullet$
node using the fact that this can be only a collection (of elements) or a primitive value.

Finally, the XML type provider in F\# Data includes an additional option to use \emph{global inference}.
In that case, the inference algorithm from Section~\ref{sec:formal-inferval} is replaced with an alternative
that unifies the types of all records with the same name. This is useful when the sample is, for example,
an XHTML document -- in this case, all occurrences of an element (\emph{e.g.}~{\small\ttfamily <a>})
are treated as the same type.

% -------------------------------------------------------------------------------------------------

\begin{figure*}
\noindent
\begin{equation*}
\inference[(col)\;]
  { \tytagof(\sigma_i) = \tytagof(\sigma'_i) \quad 
    \sigma_i \tsep \sigma'_i \vdash \sigma''_i \quad
    \phi_i \tsep \phi'_i \vdash \phi''_i \qquad (\forall i\in \{ 1 .. k \}) }
  { \begin{array}{c}
    [\sigma_1,\psi_1 | \ldots | \sigma_k,\psi_k | \ldots | \sigma_n,\psi_n ] \tsep 
    [\sigma'_1,\psi'_1 | \ldots | \sigma'_k, \psi'_k | \ldots | \sigma'_m, \psi'_m ] \vdash\\[0.1em]
    [\sigma''_1, \psi''_1 | \ldots | \sigma''_k, \psi''_k | 
        \sigma_{k+1},\addopt{\psi_{k+1}} | \ldots | \sigma_{n},\addopt{\psi_{n}} |
        \sigma'_{k+1},\addopt{\psi'_{k+1}} | \ldots | \sigma'_{m}, \addopt{\psi'_{m}} ]
    \end{array} }
\quad
\begin{array}{rcl}
 \\
 \addopt{\phi} = \phi'&\narrow{\textnormal{such that}}&\phi \tsep 1? \vdash \phi' \\[0.1em]
 \phi \tsep \phi' \vdash \phi&\narrow{\textnormal{when}}&\phi :> \phi' \\[0.1em]
 \phi \tsep \phi' \vdash \phi'&\narrow{\textnormal{when}}&\phi' :> \phi
\end{array}    
\end{equation*}
\caption{Inference judgement that defines the common supertype of two heterogeneous collections}
\label{fig:subtyping-hetcol}
\end{figure*}

% -------------------------------------------------------------------------------------------------

\subsection{Heterogeneous collections}
\label{sec:impl-collections}

When introducing the XML type provider (Section~\ref{sec:providers-xml}), we briefly mentioned 
that F\# Data implements a special handling of heterogeneous collections. In the example, the provider
generated a type with \ident{Title} member of type \ident{string} (corresponding to a nested element
{\ttfamily\small <title>} that appears exactly once) and \ident{Items} member (returning a list of
values obtained from multiple {\ttfamily\small <item>} elements).

To capture this behaviour, we need to extend our earlier treatment of collections. Rather than storing
a single type for the elements as in $[\sigma]$, we store multiple possible element types. However,
this is not just a union type as we also store \emph{inferred multiplicity} of elements for 
each of the types:
%
\begin{equation*}
\begin{array}{rcl}
 \psi &\narrow{=}& 1 \lsep 1? \lsep \ast \\
 \sigma &\narrow{=}& ~\ldots \lsep [\sigma_1, \psi_1 | \ldots | \sigma_n, \psi_n ]
\end{array}
\end{equation*}
%
The multiplicities $1, 1?$ and $\ast$ represent \emph{exactly one}, \emph{zero or one} and \emph{zero or more},
respectively. Thus, the inferred type of the collection in the RSS feed would be
$[\ident{title}~\{ \ldots \}\;,1|\;\ident{item}~\{ \ldots \},\ast]$, which reads as a collection
containing exactly one \ident{title} record and any number of \ident{item} records.

The subtyping relation on heterogeneous collections specifies that a subtype can have fewer cases 
provided that they do not have to appear exactly once. A subtype can also have a stricter specification
of multiplicity (the ordering is $\ast :> 1? :> 1$).

Finding a common supertype of heterogeneous collections is analogous to the handling of union types.
The key rule is (\emph{col}) in Figure~\ref{fig:subtyping-hetcol}. It merges cases with the same tag
(by merging both the type and the multiplicity). For cases that appear only in one collection, we ensure
that the multiplicity is \emph{zero or one} or \emph{zero or more} using the auxiliary definition
$\addopt{-}$.

% -------------------------------------------------------------------------------------------------

\subsection{Inferring and generating unions}
\label{sec:impl-unions}

The F\# Data type providers prefer types that do not contain unions (Theorem~\ref{thm:no-unions}).
The motivation for this is twofold. First, the current support for type providers in the F\# compiler
does not allow type providers to generate F\#-specific types (including discriminated unions). This
means that the provided type cannot use idiomatic F\# representation. Second, when accessing data,
using records aids the explorability -- when a value is a record, users can type ``.'' and most modern
F\# editors provide auto-completion list with members.

This choice has an important desirable consequence -- it allows treating a union with additional 
cases as a subtype of another union and, hence, the provided code is guaranteed to work on \emph{more}
inputs. Consider a type that is either a \ident{Person} record or just a string. 
The type provider exposes it as the following F\# class:
%
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{PersonOrString}~=  \\[0.1em]
 \quad \kvd{member}~\ident{Person}~:~\ident{option}\langl \ident{Person}\rangl \\[0.1em]
 \quad \kvd{member}~\ident{String}~:~\ident{option}\langl \ident{string}\rangl \\[0.1em]
\end{array}
\end{equation*}
%
If we provided an algebraic data type (discriminated union), the consumer would have to use pattern
matching that would cover \emph{two} cases. The above type forces the user to also handle a third
case, when both properties return \ident{None}. This also means that union types can silently skip
over \kvd{null} values.

That said, exposing a discriminated union (perhaps together with better tooling) is certainly an
attractive alternative that we plan to consider in the future. This can also be combined with type 
inference that chooses between records and union types based on some statistical metric (see
related work in Section~\ref{sec:related}).

% -------------------------------------------------------------------------------------------------

\subsection{Practical experiences}

An important concern about using F\# Data in practice is the handling of schema change. When using a
type provider, the sample is captured at compile-time. If the schema changes later (so that the actual
input is no longer a subtype of the samples), the program using the type provider fails at run-time
and it is the developer's responsibility to handle the exception. However, this is the same problem that
happens when reading data using any other library.

F\# Data can help discover such errors earlier. The example in Section~\ref{sec:introduction}
points the JSON type provider to a sample using a live URL. This has the advantage that a re-compilation 
fails when the schema changes, which is an indication that the program needs to be updated to reflect the
change. If this is undesirable, it is always possible to cache the sample locally.

In general, there is no better solution for plain XML, CSV and JSON data sources. Some data sources provide 
versioning support (with meta-data about how the schema changed). For those, a type 
provider could adapt automatically, but we leave this for future.

An approach that we find useful in practice is to keep a set of samples. When the program fails at run-time
(because a service returned data in an unexpected format), the new input can be added as an additional sample,
which then shows what parts of code need to be modified.

% ==================================================================================================

\section{Related and future work}
\label{sec:related}

The F\# Data library connects two lines of research that have been previously disconnected. The first is 
extending the type systems of programming languages to accommodate external data sources and the second
is inferring types for real-world data sources.

The type provider mechanism has been introduced in F\# \cite{fsharp-typeprov,fsharp-typeprov-ddfp}, 
added to Agda \cite{agda-tp} and used in areas such as semantic web \cite{liteq}. Our paper is novel 
in that it shows the programming language theory behind a concrete type providers.

\paragraph{Extending the type systems.} 
A number of systems integrate external data formats into a programming language. Those include 
XML \cite{xduce,xduce-ml} and databases \cite{links}. In both of these, the system either requires
the user to explicitly define the schema (using the host language) or it has an ad-hoc extension 
that reads the schema (\emph{e.g.}~from a database). LINQ \cite{linq} is more general, but relies
on code generation when importing the schema.

The work that is most similar to F\# Data is the XML and SQL integration in C$\omega$ \cite{comega-xs}.
It extends a C\# like object-oriented language with types similar to our structural types 
(including nullable types, choices with subtyping and heterogeneous collections with multiplicities).
However, C$\omega$ does not infer the types from samples and extends the type system of the host
language (rather than using a general purpose embedding mechanism).

\paragraph{Advanced type systems.}
Aside from type providers, a number of other advanced type system features could be used to
tackle the problem discussed in this paper. The Ur \cite{ur} language has a rich system for working
with records; meta-programming \cite{template-hask}, \cite{th-camlp4} and multi-stage programming \cite{multi-stage}
could be used to generate code for the provided types. However, as far as we are aware, none of these 
systems have been used to provide the same level of integration with XML, CSV and JSON.

Another approach would be to use gradual typing \cite{gradual,gradual-js} and add types for structured
data formats to existing dynamic language to check, for example, JSON manipulation in JavaScript.

\paragraph{Typing real-world data.}
The second line of research related to our work focuses on inferring structure of real-world data sets.
A recent work on JSON \cite{typing-json} infers a succinct type using MapReduce to handle large number
of samples. It fuses similar types based on a type similarity measure. This is more sophisticated than
our technique, but it would make formally specifying the safety properties (Theorem~\ref{thm:safety}) difficult.
Extending our \emph{relativized safety} property to a \emph{probabilistic safety} is an interesting 
future work.

The PADS project \cite{pads-dsl,pads-ml} tackles a more general problem of handling \emph{any} data format.
The schema definitions in PADS are similar to our structural type. The structure inference for LearnPADS
\cite{pads-learn} infers the data format from a flat input stream. A PADS type provider could follow
many of the patterns we explore in this paper, but formally specifying the safety property would be
challenging.

\section{Conclusions}
\label{sec:conclusions}

In this paper, we explored the F\# Data type providers for structured data formats such as XML, CSV and JSON.
As most real-world data do not come with an explicit schema, the library uses \emph{type inference} that
deduces a type from a set of samples. Next, it uses the \emph{type provider} mechanism to integrate the 
inferred type directly into the F\# type system.

The type inference algorithm we use is based on a common supertype relation. For usability reasons, the
algorithm attempts to infer type that does not contain unions and prefers records with optional fields instead.
The algorithm is simple and predictable, which is important as developers need to understand how changing the
samples affects the resulting types.

In the second part of the paper, we explore the programming language theory behind type providers. F\# Data
is a prime example of type providers, but our work also demonstrates a more general point. The types generated
by type providers can depend on external input (such as samples) and so we can only formulate \emph{relativized
safety property}, which says that a program is safe only if the actual inputs satisfy additional conditions --
in our case, they have to be subtypes of one of the samples.

The type provider mechanism has been described before, but this paper is novel in that it explores concrete
type providers from the perspective of programming language theory. This is particularly important as the
F\# Data library is becoming de-facto standard tool for data access in F\#\footnote{At the time of writing, 
the library has 37,000 downloads on NuGet; over 1600 commits and 36 contributors on GitHub.} and other 
languages are beginning to adopt mechanisms similar to F\# type providers.

\acks
?
We would like to thank to the F\# Data contributors on GitHub and other colleagues working on type providers,
including Jomo Fisher, Keith Battocchi and Kenji Takeda.

\bibliographystyle{abbrvnat}
\bibliography{paper}

\end{document}